---
title: "naive bayes penguins"
format: html
editor: visual
---

### Libraries

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(bayesrules)
library(janitor)
library(e1071)
options(scipen = 99)
```

### Story

We’ll start our naive Bayes classification with just a single penguin. Suppose an Antarctic researcher comes across a penguin that weighs less than 4200g with a 195mm-long flipper and 50mm-long bill. Our goal is to help this researcher identify the species of this penguin, Adelie, Chinstrap, or Gentoo.

#### Let's Plot!

Make a visualization that will help the researcher determine the species based on the characteristics described above

```{r}
data(penguins_bayes)
penguins <- penguins_bayes
```

```{r}
ggplot(data = penguins, aes(x = above_average_weight, fill = species)) +
  geom_bar()
```

```{r, warning=FALSE, message=FALSE}
ggplot(data = penguins, aes(bill_length_mm, flipper_length_mm, color = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(~species)
```

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = bill_length_mm, color = species)) +
  geom_point() +
  facet_wrap(~ above_average_weight) +
  theme_minimal() +
  scale_color_viridis_d() +
  geom_hline(yintercept = 50) +
  geom_vline(xintercept = 195)
```

Replace 1 with yes and 0 with no

```{r}
penguins <- penguins %>%
  mutate(above_average_weight = if_else(above_average_weight == 1, "yes", "no"))
```

Our old bayesian ways

P(B\|A)

-   P(Chinstrap \| below average weight)

```{r}
penguins %>% 
  tabyl(above_average_weight, species) %>% 
  adorn_percentages("row")
```

#### Naive Bayes Classification!

```{r}
naive_model_hints <- naiveBayes(species ~ flipper_length_mm + bill_length_mm + above_average_weight, data = penguins)
```

Our penguin

```{r}
our_penguin <- data.frame(bill_length_mm = 50, flipper_length_mm = 195, above_average_weight = "no")
```

Ask our model to make a prediction of what species our penguin is

```{r}
predict(naive_model_hints, newdata = our_penguin, type = "raw")
```

Test our model fro accuracy with confusion matrix

```{r}
penguins <- penguins %>% 
  mutate(predicted_species = predict(naive_model_hints, newdata = .))
```

```{r}
penguins %>% 
  tabyl(species, predicted_species) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

```{r}
naive_model_hints
```

## Lecture 2 - Under the hood of Naive Bayes

### Adelie

#### Species

```{r}
penguins %>% 
  tabyl(species)
```

#### Flipper Length - Probability density function

-   first argument = our penguins flipper length

-   second argument = mean flipper length for adelie penguins

-   third argument = standard deviation in flipper length for adelie penguins

```{r}
dnorm(195, mean = 190, sd = 6.54)
```

What happened here

-   a picture worth a thousand words

```{r}
# Filter penguins dataset for Adelie species
adelie_penguins <- subset(penguins, species == "Adelie")

# Calculate mean and standard deviation of flipper_length_mm for Adelie penguins
mean_flipper <- mean(adelie_penguins$flipper_length_mm, na.rm = TRUE)
sd_flipper <- sd(adelie_penguins$flipper_length_mm, na.rm = TRUE)

# Value to highlight (195 mm)
highlight_value <- 195
highlight_density <- dnorm(highlight_value, mean = mean_flipper, sd = sd_flipper)

# Plot the normal distribution with a point for 195 mm and 1 standard deviation lines
ggplot(adelie_penguins, aes(x = flipper_length_mm)) +
  geom_density(color = "blue", size = 1) +  # Plot the density curve
  
  # Highlight the point for 195 mm
  geom_point(aes(x = highlight_value, y = highlight_density), 
             color = "red", size = 3) +  
  
  # Add a vertical dashed line for 195 mm
  geom_vline(xintercept = highlight_value, linetype = "dashed", color = "red") +  
  
  # Add vertical lines for 1 standard deviation from the mean
  geom_vline(xintercept = mean_flipper + sd_flipper, linetype = "dotted", color = "black") + 
  geom_vline(xintercept = mean_flipper - sd_flipper, linetype = "dotted", color = "black") + 
  
  # Add annotations for standard deviation lines
  annotate("text", x = mean_flipper + sd_flipper + 1, y = 0.01, 
           label = "+1 SD", color = "black", hjust = 0) +
  annotate("text", x = mean_flipper - sd_flipper - 1, y = 0.01, 
           label = "-1 SD", color = "black", hjust = 1) +
  
  # Add labels and title
  labs(title = "Normal Distribution of Flipper Lengths for Adelie Penguins",
       subtitle = paste("Mean =", round(mean_flipper, 2), "mm, SD =", round(sd_flipper, 2), "mm"),
       x = "Flipper Length (mm)", y = "Density") +
  
  # Add annotation for 195 mm
  annotate("text", x = highlight_value + 2, y = highlight_density, 
           label = paste0("195 mm\nDensity: ", round(highlight_density, 5)),
           hjust = 0, color = "red") +
  
  # Minimal theme for cleaner look
  theme_minimal()
```

#### Bill Length

```{r}
dnorm(50, mean = 38.79139, sd = 2.663405)
```

#### Above Average Weight

```{r}
penguins %>% 
  tabyl(species, above_average_weight) %>% 
  adorn_percentages("row")
```

P(Adelie) \* P(Adelie \| 195 mm flipper) \* P(Adelie \| 50 mm bill) \* P(under average weight \| Adelie)

```{r}
probs_adelie <- (152/342) * 0.04554 * 0.0000212  * 0.83
probs_adelie

```

### Chinstrap

#### Species

```{r}
penguins %>% 
  tabyl(species)
```

#### Flipper Length

```{r}
dnorm(195, mean = 195.8235, sd = 7.131894)
```

#### Bill Length

```{r}
dnorm(50, mean = 48.83382, sd = 3.339256)
```

#### Above Average Weight

```{r}
penguins %>% 
  tabyl(species, above_average_weight) %>% 
  adorn_percentages("row")
```

P(Chinstrap) \* P(Chinstrap \| 195 mm flipper) \* P(Chinstrap \| 50 mm bill) \* P(under average weight \| Chinstrap)

```{r}
probs_chinstrap <- (68/342) * 0.05541 * 0.112 * 0.89
probs_chinstrap

```

### Find Gentoo
