---
title: "Test Results"
format: html
editor: visual
---

#### Load Libraries

```{r, message=FALSE}
library(tidyverse)
library(vembedr)
```

```{r}
embed_url("https://www.youtube.com/watch?v=R13BD8qKeTg")
```

#### The Story

You test positive for a rare disease that only effects 0.001 (One in one thousand people).

So you ask the doctor:

- How certain is it that I have this disease?
  - The test correctly identifies 99% of people that have the disease and only incorrectly identifies 1% of people that don't have the disease

What are the chances that you actually have this disease?

- Some would say 99%, the accuracy of the test What does bayes say?

$$P(B \mid A) = \frac{P(B) P(A \mid B)}{P(A)}$$

B \<- Has Disease

A \<- Positive test result

P(B\|A) \<- The probability of having the disease given a positive test result

#### Simulate the Data

```{r}
set.seed(70)  # For reproducibility

# Parameters
n_patients <- 10000  # Total population size
n_diseased <- 10     # Number of patients with the disease
sensitivity <- 0.99  # True positive rate (sensitivity)
false_positive_rate <- 0.01  # False positive rate

# Step 1: Create the DataFrame with patients
patients <- data.frame(
  patient_id = 1:n_patients,
  has_disease = c(rep("yes", n_diseased), rep("no", n_patients - n_diseased))  # 10 with the disease, rest without
)

# Shuffle the DataFrame to randomize patient order
patients <- patients %>%
  sample_frac(size = 1)

# Step 2: Simulate the test results based on disease status
patients <- patients %>%
  mutate(
    # Test result is positive if the person has the disease and the test is sensitive,
    # or if they don't have the disease but it's a false positive
    test_result = case_when(
      has_disease == "yes" & rbinom(n_patients, size = 1, prob = sensitivity) == 1 ~ "positive",
      has_disease == "no" & rbinom(n_patients, size = 1, prob = false_positive_rate) == 1 ~ "positive",
      TRUE ~ "negative"
    )
  )
```

#### Apply Bayes Theorem in Class

##### P(B)

```{r}
patients %>% 
  count(has_disease) %>% 
  mutate(percent = n / sum(n))
```

```{r}
probability_disease <- 0.001
```

##### P(A\|B)

```{r}
patients %>% 
  count(has_disease, test_result) %>% 
  group_by(has_disease) %>% 
  mutate(percent = n / sum(n)) 
```

```{r}
probability_positive_result_given_disease <- 1
```

##### P(A)

```{r}
patients %>% 
  count(test_result) %>% 
  mutate(percent = n / sum(n))
```

```{r}
probability_positive_test_result <- 0.0102
```

##### P(B\|A)

```{r}
(probability_positive_result_given_disease * probability_disease) / probability_positive_test_result
```

#### 
