---
title: "Bayes Regression App"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(bayesrules)
```

### Data

```{r}
bikes_data <- bayesrules::bikes
```

### Question

How many rides will we have on a 75 degree (feel) day

```{r}
bikes_data |> 
  filter(between(temp_feel, 74.9, 76)) |> 
  ggplot(aes(x = rides)) +
  geom_histogram()
```

```{r}
ggplot(bikes_data, aes(x = temp_feel, y = rides)) +
  geom_point()
```

### App

```{r}
library(shiny)
library(tidyverse)
library(bayesrules)   # provides the `bikes` data AND plot_normal()/plot_gamma()
library(rstanarm)     # for the real stan_glm() MCMC fit

bikes_data <- bayesrules::bikes
mean_temp  <- mean(bikes_data$temp_feel)
temp_range <- range(bikes_data$temp_feel)

# "Data alone" reference for beta1: a plain OLS fit, computed once since
# bikes_data never changes. Its sampling distribution (estimate +/- SE)
# is the frequentist stand-in for "what does the data alone say", the
# direct regression analogue of the scaled-likelihood curve from the
# Beta-Binomial election app.
ols_model     <- lm(rides ~ temp_feel, data = bikes_data)
ols_coef      <- summary(ols_model)$coefficients
ols_beta1_est <- unname(ols_coef["temp_feel", "Estimate"])
ols_beta1_se  <- unname(ols_coef["temp_feel", "Std. Error"])

# "Data alone" reference for beta0c (typical ridership at the average
# temperature): the OLS fitted value AT mean_temp, which -- by a basic
# property of least-squares regression -- always exactly equals
# mean(rides), with its own standard error from predict().
ols_mu_pred <- predict(ols_model, newdata = data.frame(temp_feel = mean_temp), se.fit = TRUE)
ols_mu_est  <- unname(ols_mu_pred$fit)
ols_mu_se   <- unname(ols_mu_pred$se.fit)

app_css <- tags$head(tags$style(HTML("
  .story-panel {
    background: #eef4fb; border-left: 4px solid #4C72B0;
    border-radius: 4px; padding: 10px 16px; margin-bottom: 14px;
  }
  .interp-panel {
    background: #f2f8f0; border-left: 4px solid #55A868;
    border-radius: 4px; padding: 10px 16px; margin-top: 10px;
  }
  .concept-panel {
    background: #f7f2fb; border-left: 4px solid #8B5CF6;
    border-radius: 4px; padding: 10px 16px; margin-top: 10px;
    margin-bottom: 10px; font-size: 0.95em;
  }
  .code-box {
    background: #2b2b2b; color: #eaeaea; font-family: monospace;
    font-size: 13px; padding: 12px 16px; border-radius: 6px;
    white-space: pre-wrap; margin-top: 8px;
  }
  .question-card {
    border-left: 4px solid #4C72B0;
    background: #fafafa; border-radius: 4px;
    padding: 10px 16px; margin-bottom: 12px;
  }
  .question-tab-badge {
    display: inline-block; margin-left: 8px; font-size: 0.65em;
    font-weight: 600; text-transform: uppercase; letter-spacing: 0.03em;
    color: #4C72B0; background: #eef4fb; border-radius: 10px;
    padding: 2px 9px; vertical-align: middle;
  }
")))

## ------------------------------------------------------------
## Helpers
## ------------------------------------------------------------

# Simulate directly from the priors (Exponential(l) == Gamma(1, l), so
# rexp() and bayesrules::plot_gamma(shape=1, rate=l) describe the same
# sigma prior -- this reuses the book's own plotting functions.
simulate_prior_draws <- function(m0, s0, m1, s1, l, n = 200) {
  tibble(
    beta0c = rnorm(n, m0, s0),
    beta1  = rnorm(n, m1, s1),
    sigma  = rexp(n, l)
  )
}

# beta0c is the CENTERED intercept (typical ridership at the average
# temperature) -- to draw a line against raw temp_feel we decenter it,
# exactly the beta0 vs beta0c distinction from the book's Figure 9.3.
prior_line_df <- function(draws, x_seq) {
  draws %>%
    mutate(draw_id = row_number()) %>%
    crossing(x = x_seq) %>%
    mutate(mu = beta0c + beta1 * (x - mean_temp))
}

stan_call_text <- function(m0, s0, m1, s1, l, prior_pd, chains, iter) {
  sprintf(
    paste0(
      "stan_glm(\n",
      "  rides ~ temp_feel, data = bikes,\n",
      "  family = gaussian,\n",
      "  prior_intercept = normal(%s, %s),\n",
      "  prior = normal(%s, %s),\n",
      "  prior_aux = exponential(%s),\n",
      "  chains = %d, iter = %d, seed = 84735%s)"
    ),
    m0, s0, m1, s1, l, chains, iter,
    if (prior_pd) ",\n  prior_PD = TRUE" else ""
  )
}

# One reference card per question: a tab badge and the prompt, no
# answer box -- answers go in the comment block at the top of this
# script instead, so they persist with the file rather than the app.
question_card <- function(n, tab, prompt) {
  div(class = "question-card",
      h5(sprintf("Question %d", n), tags$span(class = "question-tab-badge", tab)),
      p(prompt)
  )
}

## ------------------------------------------------------------
## UI
## ------------------------------------------------------------

ui <- navbarPage(
  title = "Bayesian Bike Regression",
  header = app_css,
  
  ## ---------------- TAB 1: Tune the prior ----------------
  tabPanel(
    "1. Tune the Prior",
    sidebarLayout(
      sidebarPanel(
        div(class = "story-panel",
            strong("The three unknowns:"), " beta0c (typical ridership at ",
            "the average temperature), beta1 (change in ridership per ",
            "degree), and sigma (day-to-day scatter around the line). ",
            "Drag the sliders and watch the plots and the code below ",
            "update together."
        ),
        h4("beta0c ~ Normal(m0, s0)"),
        sliderInput("m0", "m0 (typical ridership at avg. temp)", 0, 10000, 5000, step = 100),
        sliderInput("s0", "s0 (uncertainty around that typical value)", 100, 3000, 1000, step = 50),
        h4("beta1 ~ Normal(m1, s1)"),
        sliderInput("m1", "m1 (typical change per degree)", -50, 300, 100, step = 5),
        sliderInput("s1", "s1 (uncertainty around that change)", 5, 150, 40, step = 5),
        h4("sigma ~ Exponential(l)"),
        sliderInput("l", "l (rate; E(sigma) = 1/l)", 0.0002, 0.005, 0.0008, step = 0.0001),
        actionButton("reset_prior", "Reset to the text's priors"),
        div(class = "concept-panel",
            strong("Reading a 95% range from mean and SD:"), " for a Normal ",
            "distribution, 95% of the plausible values fall within about ",
            "mean \u00b1 1.96\u00d7SD -- not \u00b1SD (that's only the 68% range). ",
            "So Normal(100, 40) puts 95% of its mass between roughly ",
            "100 - 1.96(40) = 22 and 100 + 1.96(40) = 178, not the 60-140 ",
            "you'd get from just \u00b1SD."
        ),
        div(class = "concept-panel",
            strong("Why sigma's curve looks so different:"), " beta0c and ",
            "beta1 use Normal priors, which are symmetric -- their mean, ",
            "median, and peak (mode) all sit at the same spot. sigma's ",
            "Exponential prior is NOT symmetric: it peaks at 0, but its ",
            tags$em("mean"), " (E(sigma) = 1/l) sits well to the right of ",
            "that peak, pulled there by a long thin tail of larger values. ",
            "The peak tells you the single most likely value; the mean is ",
            "a full weighted average across the whole tail. They're both ",
            "correct descriptions of the same curve -- they're just ",
            "answering different questions (\"what's most likely?\" vs. ",
            "\"what's the long-run average?\")."
        ),
        div(class = "concept-panel",
            strong("Why sigma can't go negative:"), " sigma measures ",
            "spread -- how far typical days scatter from the average. ",
            "Spread can't be negative any more than a distance can. The ",
            "Exponential formula itself enforces this: it's only defined ",
            "for sigma \u2265 0. Normal priors, by contrast, are fine for ",
            "beta0c and beta1 because those parameters really can be any ",
            "real number -- including 0 or negative."
        )
      ),
      mainPanel(
        uiOutput("priorStatement"),
        fluidRow(
          column(4, plotOutput("beta0Plot", height = "220px")),
          column(4, plotOutput("beta1Plot", height = "220px")),
          column(4, plotOutput("sigmaPlot", height = "220px"))
        ),
        h4("200 prior plausible model lines"),
        plotOutput("priorLinesPlot", height = "280px"),
        div(class = "concept-panel",
            strong("Why the lines funnel together in the middle:"), " beta0c ",
            "is defined as ridership at the ", tags$em("average"), " temperature ",
            sprintf("(~%.0f\u00b0F), not at 0\u00b0F", mean_temp), " -- so every ",
            "line's uncertainty is smallest right there and grows the further ",
            "you move toward either edge, since a small disagreement about ",
            "the slope compounds over distance. A few lines dipping below 0 ",
            "rides at the cold end isn't a data problem -- it's a property ",
            "of the prior itself: Normal distributions don't know negative ",
            "ridership is impossible."
        ),
        h4("Equivalent stan_glm() code (prior only, prior_PD = TRUE)"),
        div(class = "code-box", textOutput("priorCode"))
      )
    )
  ),
  
  ## ---------------- TAB 2: Fit the posterior ----------------
  tabPanel(
    "2. Fit the Posterior",
    sidebarLayout(
      sidebarPanel(
        div(class = "story-panel",
            strong("This runs real MCMC"), " on the actual ", code("bikes"),
            " dataset (500 days), using the priors you just tuned on Tab 1. ",
            "Unlike Tab 1's instant prior simulation, this genuinely samples ",
            "the posterior with Stan -- expect it to take a few seconds."
        ),
        p(strong("Current priors (from Tab 1):")),
        verbatimTextOutput("currentPriorSummary"),
        actionButton("fit_posterior", "Fit the posterior model", class = "btn-primary"),
        hr(),
        helpText("Uses 4 chains, 2000 iterations each (fewer than the ",
                 "book's 10,000 to keep this responsive in a live demo -- ",
                 "more iterations would only reduce simulation noise, not ",
                 "change the substantive answer, since the real driver of ",
                 "precision here is the 500 real data points, not the ",
                 "chain length.")
      ),
      mainPanel(
        uiOutput("posteriorStatus"),
        conditionalPanel(
          condition = "output.modelFitted == true",
          h4("Posterior summary"),
          tableOutput("posteriorTable"),
          uiOutput("posteriorInterpretation"),
          h4("Prior belief vs. data alone vs. posterior, for beta1 specifically"),
          div(class = "concept-panel",
              strong("This is the plot Q5 is really asking for:"), " three ",
              "separate estimates of \"expected change in rides per 1-degree ",
              "increase\" side by side -- your prior belief (m1, before any ",
              "data), what the data alone says (a plain OLS fit, ignoring the ",
              "prior entirely), and the posterior (the balance between them). ",
              "This is the direct regression version of the prior/likelihood/",
              "posterior plot from the earlier Beta-Binomial election example."
          ),
          tableOutput("beta1CompareTable"),
          plotOutput("beta1ComparePlot", height = "280px"),
          div(class = "concept-panel",
              strong("watch the SE:"), " the data-alone curve is built from ",
              sprintf("OLS's estimate (%.1f) and its standard error (%.2f).",
                      ols_beta1_est, ols_beta1_se),
              " With n = 500 real days, that standard error is small, so the ",
              "data-alone curve is naturally tall and narrow -- a sign of how ",
              "precisely 500 real observations pin down the slope, compared ",
              "to how spread out your prior belief (s1) was before seeing ",
              "any of them."
          ),
          h4("Prior belief vs. data alone vs. posterior, for beta0c (typical ridership at the average temperature)"),
          div(class = "concept-panel",
              strong("This is the plot Q10 needs:"), " m0 and s0 set your ",
              "prior belief about beta0c specifically, not the raw ",
              "(Intercept) in the table above (that's ridership at a ",
              "hypothetical 0\u00b0F -- a different, less meaningful number). ",
              "Try setting m0 to 10,000 (its max) on Tab 1, then refit here: ",
              "watch whether the green posterior curve follows the blue ",
              "prior curve toward 10,000, or stays anchored near the tall ",
              "orange data-alone curve instead."
          ),
          tableOutput("beta0cCompareTable"),
          plotOutput("beta0cComparePlot", height = "280px"),
          div(class = "concept-panel",
              strong("Why the posterior should land close to the textbook's numbers:"),
              " with Normal priors, there isn't a single clean number like ",
              "\"alpha + beta\" to compare against n the way there is for a ",
              "Beta prior -- but the same core idea still applies: a prior's ",
              "pull on the posterior depends on how tightly it's specified ",
              "(small s0/s1 = a more confident, more stubborn prior) relative ",
              "to how much real data you have. With n = 500 real days and ",
              "priors that are informative but not extremely narrow (s1 = 40, ",
              "not s1 = 2), the data comfortably dominates -- so a posterior ",
              "landing close to what plain OLS would give you isn't a ",
              "coincidence, it's exactly what you'd expect."
          ),
          div(class = "concept-panel",
              strong("Reading \"P(beta1 > 0 | data)\":"), " this is computed by ",
              "literally counting what fraction of the thousands of posterior ",
              "draws are positive -- a direct probability statement about the ",
              "parameter itself. A p-value, by contrast, answers a different ",
              "question: \"if the true effect were exactly zero, how surprising ",
              "would this data be?\" Both can point the same direction, but ",
              "only the Bayesian probability is allowed to say \"there's an X% ",
              "chance the effect is positive.\""
          ),
          h4("50 posterior plausible model lines vs. the real data"),
          plotOutput("posteriorLinesPlot", height = "300px"),
          h4("Posterior distribution of beta1 (temp_feel effect)"),
          plotOutput("beta1PosteriorPlot", height = "220px"),
          h4("Equivalent stan_glm() code (this is the exact call used)"),
          div(class = "code-box", textOutput("posteriorCode")),
          h4("MCMC diagnostics"),
          verbatimTextOutput("diagnostics"),
          div(class = "concept-panel",
              strong("Reading Rhat and n_eff:"), " Rhat compares your several ",
              "MCMC chains against each other -- if they've all genuinely ",
              "converged on the same posterior, Rhat should sit right around ",
              "1.00 (matching what the textbook reports: 0.9999-1.0000). A ",
              "value noticeably above 1 (a common rule of thumb: worry above ",
              "roughly 1.01-1.05) means the chains disagree with each other, ",
              "which is a red flag that the simulation hasn't settled yet -- ",
              "not a statement about whether you have enough real data. ",
              "n_eff (effective sample size) tells you, roughly, how many ",
              "genuinely independent-looking draws your MCMC chain produced; ",
              "MCMC draws are naturally a bit correlated step-to-step, so ",
              "n_eff is usually somewhat smaller than the raw number of ",
              "iterations you asked for."
          )
        )
      )
    )
  ),
  
  ## ---------------- TAB 3: Posterior prediction ----------------
  tabPanel(
    "3. Predict Ridership",
    sidebarLayout(
      sidebarPanel(
        div(class = "story-panel",
            strong("Requires a fitted model from Tab 2 first."), " Once ",
            "fitted, this tab is fully live -- no new MCMC needed, since ",
            "we're just reusing the posterior draws already sitting in memory."
        ),
        sliderInput("temp_predict", "Temperature to predict for (\u00b0F)",
                    min = 20, max = 100, value = 75, step = 1),
        helpText(sprintf("The observed data ranges from %.0f\u00b0F to %.0f\u00b0F -- ",
                         temp_range[1], temp_range[2]),
                 "predicting outside that range leans more heavily on the ",
                 "model's assumed linear shape than on nearby real data.")
      ),
      mainPanel(
        uiOutput("predictionStatus"),
        conditionalPanel(
          condition = "output.modelFitted == true",
          uiOutput("predictionStatement"),
          div(class = "concept-panel",
              strong("mu vs. the posterior predictive -- two different questions:"),
              " mu asks \"what's the average ridership across ", tags$em("all"),
              " days at this temperature?\" -- it's narrow because averaging ",
              "smooths out day-to-day noise. The posterior predictive asks ",
              "\"what will happen on ", tags$em("one specific"), " day at this ",
              "temperature?\" -- it has to add real day-to-day scatter (sigma) ",
              "on top of the same uncertainty mu already has, which is why the ",
              "orange curve is always wider than the blue one."
          ),
          div(class = "concept-panel",
              strong("A real limitation, visible at cold temperatures:"), " a ",
              "Normal distribution technically allows any real number, ",
              "including negative ones -- so at cold temperatures, where ",
              "typical ridership is already low, the posterior predictive's ",
              "left tail can dip below 0 rides. That's not a bug in the ",
              "simulation; it's an honest sign that a plain Normal model is ",
              "only an approximation, one that strains more at the extremes ",
              "than in the middle of the observed data."
          ),
          plotOutput("predictionPlot", height = "280px"),
          h4("The by-hand simulation code (what's actually happening)"),
          div(class = "code-box", textOutput("predictionCode")),
          h4("The rstanarm shortcut (same answer, one line)"),
          div(class = "code-box", textOutput("shortcutCode"))
        )
      )
    )
  ),
  
  ## ---------------- TAB 4: Questions ----------------
  tabPanel(
    "4. Questions",
    fluidPage(
      div(class = "story-panel",
          strong("How to use this tab:"), " Each question names which tab ",
          "to work in. Set the sliders, read off the app's numbers, then ",
          "write your answer as a comment at the top of ", code("app.R"),
          " (the script you're running this from) -- that way your ",
          "answers are saved with the script itself, not lost when you ",
          "close the app."
      ),
      question_card(1, "Tab 1", "Set m1 = 100, s1 = 40 (the book's defaults). By hand, using mean +/- 1.96*SD, calculate the 95% plausible range for beta1. Check your answer against what the app reports."),
      question_card(2, "Tab 1", "Increase s1 from 40 to 80, keeping m1 = 100. What happens to the spread of the \"200 prior plausible lines\" plot? Explain in one sentence why."),
      question_card(3, "Tab 1", "Set l (sigma's rate) to 0.002, then to 0.0003. Compare the two sigma prior plots. Which one reflects a stronger belief that ridership is tightly predicted by temperature alone?"),
      question_card(4, "Tab 1", "The sigma prior plot peaks at 0, not in the middle like the beta0c and beta1 plots. Why? (Hint: what values is sigma even allowed to take?)"),
      question_card(5, "Tab 2", "Reset to the book's default priors and fit the posterior. Report the posterior median for the Intercept and for temp_feel. How closely do they match the textbook's own values (-2194.24 and 82.16)? Given how large n = 500 is relative to this prior's strength, is that closeness surprising, or exactly what you'd expect? Explain using the weight-of-evidence idea."),
      question_card(6, "Tab 2", "What does \"P(beta1 > 0 | data)\" actually tell you, and why is this a genuine probability statement rather than a p-value?"),
      question_card(7, "Tab 2", "Look at the MCMC diagnostics table (Rhat, n_eff). What would an Rhat far from 1 (say, 1.5) suggest about the chains?"),
      question_card(8, "Tab 3", "Predict ridership at 75\u00b0F. Record both credible intervals shown: \"typical ridership (mu)\" and \"one day's ridership (posterior predictive).\" Which is wider, and why?"),
      question_card(9, "Tab 3", "Now predict at a much colder temperature, like 30\u00b0F. Does the posterior predictive interval's lower bound dip below zero? What limitation of using a Normal distribution for ridership does that reveal?"),
      question_card(10, "Tab 1 + Tab 2, extends Q5", "The m0 slider only goes up to 10,000. In Q5 you fit the posterior with the default m0 = 5,000 and found the posterior intercept near -2194. Now set m0 to its max, 10,000 -- double your original prior belief -- keep everything else the same, and refit. Use the \"Prior belief vs. data alone vs. posterior, for beta0c\" plot to see it directly: does the posterior follow the prior toward 10,000, or stay anchored near the data-alone curve? Explain using weight-of-evidence: compare how confidently you specified m0 (via s0) to how much real data you have (n = 500).")
    )
  )
)

## ------------------------------------------------------------
## Server
## ------------------------------------------------------------

server <- function(input, output, session) {
  
  ## ===== TAB 1: Prior tuning (instant, no MCMC) =====
  
  observeEvent(input$reset_prior, {
    updateSliderInput(session, "m0", value = 5000)
    updateSliderInput(session, "s0", value = 1000)
    updateSliderInput(session, "m1", value = 100)
    updateSliderInput(session, "s1", value = 40)
    updateSliderInput(session, "l", value = 0.0008)
  })
  
  prior_draws <- reactive({
    simulate_prior_draws(input$m0, input$s0, input$m1, input$s1, input$l, n = 200)
  })
  
  output$priorStatement <- renderUI({
    m0 <- input$m0; s0 <- input$s0; m1 <- input$m1; s1 <- input$s1; l <- input$l
    div(class = "interp-panel",
        sprintf(
          paste0(
            "Before seeing any data: on a typical (~%.0f\u00b0F) day, you expect ",
            "about %.0f riders (95%% plausible range roughly %.0f to %.0f). For ",
            "every one-degree increase in temperature, you expect ridership to ",
            "change by about %.0f rides (95%% plausible range roughly %.0f to ",
            "%.0f). Day-to-day scatter around that average is expected to be ",
            "about %.0f rides (E(sigma) = 1/l)."
          ),
          mean_temp, m0, m0 - 1.96*s0, m0 + 1.96*s0,
          m1, m1 - 1.96*s1, m1 + 1.96*s1,
          1/l
        )
    )
  })
  
  output$beta0Plot <- renderPlot({
    plot_normal(mean = input$m0, sd = input$s0) + labs(x = "beta0c", y = "pdf") + theme_minimal()
  })
  output$beta1Plot <- renderPlot({
    plot_normal(mean = input$m1, sd = input$s1) + labs(x = "beta1", y = "pdf") + theme_minimal()
  })
  output$sigmaPlot <- renderPlot({
    plot_gamma(shape = 1, rate = input$l) + labs(x = "sigma", y = "pdf") + theme_minimal()
  })
  
  output$priorLinesPlot <- renderPlot({
    x_seq <- seq(temp_range[1], temp_range[2], length.out = 50)
    prior_line_df(prior_draws(), x_seq) %>%
      ggplot(aes(x = x, y = mu, group = draw_id)) +
      geom_line(alpha = 0.15, color = "#4C72B0") +
      labs(x = "temp_feel", y = "rides",
           title = "200 prior plausible relationships between temp_feel and rides") +
      theme_minimal(base_size = 13)
  })
  
  output$priorCode <- renderText({
    stan_call_text(input$m0, input$s0, input$m1, input$s1, input$l,
                   prior_pd = TRUE, chains = 4, iter = 2000)
  })
  
  ## ===== TAB 2: Real posterior fit (MCMC, behind a button) =====
  
  output$currentPriorSummary <- renderText({
    sprintf("beta0c ~ Normal(%s, %s)\nbeta1  ~ Normal(%s, %s)\nsigma  ~ Exponential(%s)",
            input$m0, input$s0, input$m1, input$s1, input$l)
  })
  
  fitted_model <- reactiveVal(NULL)
  fitted_draws <- reactiveVal(NULL)
  
  observeEvent(input$fit_posterior, {
    withProgress(message = "Running MCMC (stan_glm)...", value = 0.3, {
      m <- stan_glm(
        rides ~ temp_feel, data = bikes_data, family = gaussian,
        prior_intercept = normal(input$m0, input$s0),
        prior = normal(input$m1, input$s1),
        prior_aux = exponential(input$l),
        chains = 4, iter = 2000, seed = 84735, refresh = 0
      )
      incProgress(0.6)
      fitted_model(m)
      fitted_draws(as.data.frame(m))
    })
  })
  
  output$modelFitted <- reactive({ !is.null(fitted_model()) })
  outputOptions(output, "modelFitted", suspendWhenHidden = FALSE)
  
  output$posteriorStatus <- renderUI({
    if (is.null(fitted_model())) {
      p(em("Click \"Fit the posterior model\" to run MCMC on the real bikes data."))
    } else {
      NULL
    }
  })
  
  output$posteriorTable <- renderTable({
    req(fitted_draws())
    fitted_draws() %>%
      pivot_longer(everything(), names_to = "parameter", values_to = "value") %>%
      group_by(parameter) %>%
      summarize(
        median = median(value),
        `2.5%`  = quantile(value, 0.025),
        `97.5%` = quantile(value, 0.975),
        .groups = "drop"
      )
  }, digits = 2)
  
  output$posteriorInterpretation <- renderUI({
    req(fitted_draws())
    df <- fitted_draws()
    p_pos <- mean(df$temp_feel > 0)
    div(class = "interp-panel",
        sprintf(
          paste0(
            "Posterior median relationship: rides = %.1f + %.1f x temp_feel. ",
            "P(beta1 > 0 | data) \u2248 %.3f -- based on the real data, we're ",
            "essentially certain ridership increases with temperature."
          ),
          median(df$`(Intercept)`), median(df$temp_feel), p_pos
        )
    )
  })
  
  output$beta1CompareTable <- renderTable({
    req(fitted_draws())
    tibble(
      Source = c("Prior belief (m1)", "Data alone (OLS estimate)", "Posterior (median)"),
      `beta1 (rides per degree)` = c(
        input$m1,
        ols_beta1_est,
        median(fitted_draws()$temp_feel)
      )
    )
  }, digits = 2)
  
  output$beta1ComparePlot <- renderPlot({
    req(fitted_draws())
    post_beta1 <- fitted_draws()$temp_feel
    
    x_lo <- min(input$m1 - 3*input$s1, ols_beta1_est - 4*ols_beta1_se, min(post_beta1))
    x_hi <- max(input$m1 + 3*input$s1, ols_beta1_est + 4*ols_beta1_se, max(post_beta1))
    x_seq <- seq(x_lo, x_hi, length.out = 500)
    
    prior_df <- tibble(x = x_seq, density = dnorm(x_seq, input$m1, input$s1), curve = "Prior belief")
    data_df  <- tibble(x = x_seq, density = dnorm(x_seq, ols_beta1_est, ols_beta1_se), curve = "Data alone (OLS)")
    post_dens <- density(post_beta1, n = 500)
    post_df  <- tibble(x = post_dens$x, density = post_dens$y, curve = "Posterior")
    
    all_df <- bind_rows(prior_df, data_df, post_df) %>%
      mutate(curve = factor(curve, levels = c("Prior belief", "Data alone (OLS)", "Posterior")))
    
    ggplot(all_df, aes(x = x, y = density, color = curve, fill = curve)) +
      geom_area(alpha = 0.25, position = "identity") +
      geom_line(linewidth = 1.1) +
      scale_color_manual(values = c("Prior belief" = "#4C72B0",
                                    "Data alone (OLS)" = "#DD8452",
                                    "Posterior" = "#55A868")) +
      scale_fill_manual(values = c("Prior belief" = "#4C72B0",
                                   "Data alone (OLS)" = "#DD8452",
                                   "Posterior" = "#55A868")) +
      labs(x = "beta1 (expected change in rides per 1-degree increase)",
           y = "density", color = NULL, fill = NULL) +
      theme_minimal(base_size = 13) +
      theme(legend.position = "top")
  })
  
  output$beta0cCompareTable <- renderTable({
    req(fitted_draws())
    post_beta0c <- fitted_draws()$`(Intercept)` + fitted_draws()$temp_feel * mean_temp
    tibble(
      Source = c("Prior belief (m0)", "Data alone (OLS, at avg. temp)", "Posterior (median)"),
      `beta0c (typical rides at avg. temp)` = c(
        input$m0,
        ols_mu_est,
        median(post_beta0c)
      )
    )
  }, digits = 1)
  
  output$beta0cComparePlot <- renderPlot({
    req(fitted_draws())
    post_beta0c <- fitted_draws()$`(Intercept)` + fitted_draws()$temp_feel * mean_temp
    
    x_lo <- min(input$m0 - 3*input$s0, ols_mu_est - 4*ols_mu_se, min(post_beta0c))
    x_hi <- max(input$m0 + 3*input$s0, ols_mu_est + 4*ols_mu_se, max(post_beta0c))
    x_seq <- seq(x_lo, x_hi, length.out = 500)
    
    prior_df <- tibble(x = x_seq, density = dnorm(x_seq, input$m0, input$s0), curve = "Prior belief")
    data_df  <- tibble(x = x_seq, density = dnorm(x_seq, ols_mu_est, ols_mu_se), curve = "Data alone (OLS)")
    post_dens <- density(post_beta0c, n = 500)
    post_df  <- tibble(x = post_dens$x, density = post_dens$y, curve = "Posterior")
    
    all_df <- bind_rows(prior_df, data_df, post_df) %>%
      mutate(curve = factor(curve, levels = c("Prior belief", "Data alone (OLS)", "Posterior")))
    
    ggplot(all_df, aes(x = x, y = density, color = curve, fill = curve)) +
      geom_area(alpha = 0.25, position = "identity") +
      geom_line(linewidth = 1.1) +
      scale_color_manual(values = c("Prior belief" = "#4C72B0",
                                    "Data alone (OLS)" = "#DD8452",
                                    "Posterior" = "#55A868")) +
      scale_fill_manual(values = c("Prior belief" = "#4C72B0",
                                   "Data alone (OLS)" = "#DD8452",
                                   "Posterior" = "#55A868")) +
      labs(x = "beta0c (typical ridership at the average temperature)",
           y = "density", color = NULL, fill = NULL) +
      theme_minimal(base_size = 13) +
      theme(legend.position = "top")
  })
  
  output$posteriorLinesPlot <- renderPlot({
    req(fitted_draws())
    df <- fitted_draws() %>% slice_sample(n = 50) %>% mutate(draw_id = row_number())
    x_seq <- seq(temp_range[1], temp_range[2], length.out = 50)
    line_df <- df %>%
      crossing(x = x_seq) %>%
      mutate(mu = `(Intercept)` + temp_feel * x)
    
    ggplot() +
      geom_point(data = bikes_data, aes(x = temp_feel, y = rides), size = 0.4, alpha = 0.4) +
      geom_line(data = line_df, aes(x = x, y = mu, group = draw_id),
                color = "#55A868", alpha = 0.3) +
      labs(x = "temp_feel", y = "rides",
           title = "50 posterior plausible lines over the real observed data") +
      theme_minimal(base_size = 13)
  })
  
  output$beta1PosteriorPlot <- renderPlot({
    req(fitted_draws())
    fitted_draws() %>%
      ggplot(aes(x = temp_feel)) +
      geom_density(fill = "#55A868", alpha = 0.4) +
      geom_vline(xintercept = 0, linetype = "dashed", color = "#8B3A3A") +
      labs(x = "beta1 (posterior draws)", y = "density",
           title = "Posterior distribution of the temp_feel effect") +
      theme_minimal(base_size = 13)
  })
  
  output$posteriorCode <- renderText({
    stan_call_text(input$m0, input$s0, input$m1, input$s1, input$l,
                   prior_pd = FALSE, chains = 4, iter = 2000)
  })
  
  output$diagnostics <- renderPrint({
    req(fitted_model())
    summ <- summary(fitted_model())
    # summary() on a stanreg object already includes Rhat and n_eff columns
    # for every parameter -- no separate rhat()/neff_ratio() call needed,
    # which sidesteps a real gotcha: the `posterior` package also defines
    # a function called rhat(), but it expects a plain numeric draws array,
    # not a fitted model object. If both rstanarm and posterior are loaded,
    # whichever was loaded LAST wins the name "rhat" -- and if posterior's
    # version gets called on a stanreg object by mistake, it fails with
    # exactly the "non-numeric argument to binary operator" error, since
    # it's trying to do arithmetic on something that isn't a plain numeric
    # array. summary() avoids this ambiguity entirely.
    print(summ[, c("Rhat", "n_eff")])
  })
  
  ## ===== TAB 3: Posterior prediction (fast, reuses fitted_draws) =====
  
  output$predictionStatus <- renderUI({
    if (is.null(fitted_draws())) {
      p(em("Fit a posterior model on Tab 2 first -- this tab reuses those draws."))
    } else {
      NULL
    }
  })
  
  prediction_draws <- reactive({
    req(fitted_draws())
    df <- fitted_draws()
    mu <- df$`(Intercept)` + df$temp_feel * input$temp_predict
    y_new <- rnorm(length(mu), mean = mu, sd = df$sigma)
    tibble(mu = mu, y_new = y_new)
  })
  
  output$predictionStatement <- renderUI({
    req(prediction_draws())
    pd <- prediction_draws()
    mu_ci <- quantile(pd$mu, c(0.025, 0.975))
    y_ci  <- quantile(pd$y_new, c(0.025, 0.975))
    div(class = "interp-panel",
        sprintf(
          paste0(
            "At %.0f\u00b0F: the TYPICAL ridership across all days at this ",
            "temperature is estimated at %.0f riders (95%% credible interval ",
            "%.0f to %.0f -- narrow, since it's an average). But ridership on ",
            "any ONE specific day at %.0f\u00b0F could range much more widely: ",
            "95%% posterior predictive interval %.0f to %.0f."
          ),
          input$temp_predict, mean(pd$mu), mu_ci[1], mu_ci[2],
          input$temp_predict, y_ci[1], y_ci[2]
        )
    )
  })
  
  output$predictionPlot <- renderPlot({
    req(prediction_draws())
    pd <- prediction_draws() %>%
      pivot_longer(everything(), names_to = "quantity", values_to = "value") %>%
      mutate(quantity = recode(quantity,
                               mu = "Typical ridership (mu)",
                               y_new = "One day's ridership (posterior predictive)"))
    ggplot(pd, aes(x = value, fill = quantity)) +
      geom_density(alpha = 0.5) +
      scale_fill_manual(values = c("Typical ridership (mu)" = "#4C72B0",
                                   "One day's ridership (posterior predictive)" = "#DD8452")) +
      labs(x = "rides", y = "density", fill = NULL,
           title = sprintf("Posterior vs. posterior predictive at %.0f\u00b0F", input$temp_predict)) +
      theme_minimal(base_size = 13) +
      theme(legend.position = "top")
  })
  
  output$predictionCode <- renderText({
    sprintf(
      paste0(
        "posterior <- as.data.frame(bike_model)\n",
        "mu <- posterior$`(Intercept)` + posterior$temp_feel * %d\n",
        "y_new <- rnorm(length(mu), mean = mu, sd = posterior$sigma)"
      ),
      input$temp_predict
    )
  })
  
  output$shortcutCode <- renderText({
    sprintf(
      "posterior_predict(bike_model, newdata = data.frame(temp_feel = %d))",
      input$temp_predict
    )
  })
}

shinyApp(ui, server)

```

### Questions

1.  100−1.96(40)=100−78.4≈22 100+1.96(40)=100+78.4≈178

- 95% corresponds to mean±1.96 SD, *not* ±1 SD

  - `plot_normal(100,40)`

2.  The fan of 200 lines widens substantially, especially toward the edges of the temperature range, while staying pinched together near the average temperature
    - with s1=80, the 95% range for beta1 itself widens to roughly 100 ± 1.96(80) = -57 to 257 — meaning some of the 200 simulated lines will now even slope *downward*, which wasn't plausible at s1=40.
3.  l=0.002 represents the stronger belief in a tight relationship

- We are saying σ = 1/l. σ is the variance in rides scattered around the mean

  - 1/0.002 = 500. In other words, once temperature is accounted for, relatively small scatter remains (500 rides)

  - 1/0.003 =3,333. In other words, once temperature is accounted for, very sginificant scatter remains (3,333 rides) implying temperature is only a loose, weak predictor.

4.  sigma is restricted to non-negative values, while beta0c and beta1 can genuinely take any real number — including zero or negative. That constraint forces sigma into a fundamentally different *shape* of distribution than the other two.

- sigma is describing the scatter around the mean when temperature is controlled for, which cannot be a negative number. we could have negative slope (rides could decrease with increased temperature

5.  posterior meidan for intercept = -2,197 (not helpful because we dont see any 0 degree days in our data). posterior median for temp_feel = 82.2

- The posterior almost exactly matches the prior suggesting our 500 data points carries more weight than our relatively loos prior.

6.  This is computed by literally tallying what fraction of the thousands of MCMC draws of beta1 came out positive — a direct empirical estimate of the actual probability that the true slope exceeds zero, given this prior and this data. It's a genuine probability statement about the *parameter*. A p-value answers a structurally different question: *"if the true effect were exactly zero, how surprising would data like this be?"* — a probability about the *data*, conditional on an assumption about the parameter, never a probability about the parameter itself.
7.  it means the several MCMC chains have **not** converged to the same posterior — they're disagreeing substantially with each other.
8.  **mu's 95% credible interval**: roughly 3843 to 4095 — narrow, about 250 rides wide. **Posterior predictive 95% interval**: roughly 1500 to 6482 — about 5000 rides wide, roughly **20× wider**.
9.  **Why**: mu's interval only reflects uncertainty about the *average* relationship (from not knowing beta0/beta1 exactly) — and averaging across many days smooths out individual noise, so that uncertainty shrinks a lot with n=500. The posterior predictive interval has to add a second, much bigger source of uncertainty on top: genuine day-to-day scatter (σ≈1281), representing how much any *one specific* day's ridership naturally varies even if the true average were known perfectly.
10. even doubling m0 (a substantially "wrong," confidently-stated prior belief) should still get pulled back close to the Q5 baseline once real data enters, since n=500 comfortably outweighs a moderately-confident prior

3\.

4\.

5\.

6\.

7\.

8\.

9\.

10\.
