---
title: "Fake News"
format: html
editor: visual
---

### Goal

How can we model whether news articles are fake or not with our bayesian toolkit

```{r}
library(tidyverse)
library(bayesrules)
library(janitor)
library(e1071)
library(tidytext)
```

### Data

```{r}
?bayesrules::fake_news
```

```{r}
news_data <- bayesrules::fake_news |> 
  mutate(id = row_number())
```

### Explore

```{r}

```

```{r}

```

```{r}

```

### Text-Specific Viz

```{r}
text_tokens_clean <- news_data %>%
  select(id, type, text) %>%
  mutate(text = str_replace_all(text, "['’]s\\b", "")) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words, by = "word")
```

```{r}
library(ggwordcloud)

text_tokens_clean %>%
  count(type, word, sort = TRUE) %>%
  group_by(type) %>%
  slice_max(n, n = 60) %>%
  ungroup() %>%
  ggplot(aes(label = word, size = n, color = type)) +
  geom_text_wordcloud(rm_outside = TRUE) +
  scale_size_area(max_size = 12) +
  scale_color_manual(values = c(fake = "#C0392B", real = "#27824C")) +
  facet_wrap(~type) +
  theme_minimal()
```

```{r}
tf_idf_by_type <- text_tokens_clean %>%
  count(type, word) %>%
  bind_tf_idf(word, type, n) %>%
  arrange(desc(tf_idf))

tf_idf_by_type %>%
  group_by(type) %>%
  slice_max(tf_idf, n = 15) %>%
  ungroup() %>%
  ggplot(aes(x = reorder_within(word, tf_idf, type), y = tf_idf, fill = type)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_manual(values = c(fake = "#C0392B", real = "#27824C")) +
  facet_wrap(~type, scales = "free_y") +
  labs(x = NULL, y = "TF-IDF", title = "Most distinctive words by article type") +
  theme_minimal()
```

```{r}
bigrams <- news_data %>%
  select(id, type, text) %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word, !word2 %in% stop_words$word) %>%
  unite(bigram, word1, word2, sep = " ")

bigrams %>%
  count(type, bigram, sort = TRUE) %>%
  group_by(type) %>%
  slice_max(n, n = 12) %>%
  ungroup() %>%
  ggplot(aes(x = reorder_within(bigram, n, type), y = n, fill = type)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_manual(values = c(fake = "#C0392B", real = "#27824C")) +
  facet_wrap(~type, scales = "free_y") +
  labs(x = NULL, y = "Count", title = "Most common two-word phrases by article type") +
  theme_minimal()
```

```{r}
library(igraph)
library(ggraph)

# Reuse the bigram structure, but keep both words and count co-occurrences
bigram_counts <- news_data %>%
  select(id, type, text) %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word, !word2 %in% stop_words$word,
         !is.na(word1), !is.na(word2))

# Build and plot one network, filtering to bigrams that occur often enough
# to be worth drawing -- otherwise the graph becomes an unreadable tangle
plot_bigram_network <- function(data, article_type, min_count = 3, title_color) {
  graph_data <- data %>%
    filter(type == article_type) %>%
    count(word1, word2, sort = TRUE) %>%
    filter(n >= min_count)
  
  g <- graph_from_data_frame(graph_data)
  
  ggraph(g, layout = "fr") +
    geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                   arrow = arrow(length = unit(2, "mm")), end_cap = circle(2, "mm")) +
    geom_node_point(color = title_color, size = 3) +
    geom_node_text(aes(label = name), vjust = 1.3, hjust = 1, size = 3) +
    labs(title = paste("Word co-occurrence network:", article_type)) +
    theme_void()
}

plot_bigram_network(bigram_counts, "fake", min_count = 3, title_color = "#C0392B")
plot_bigram_network(bigram_counts, "real", min_count = 3, title_color = "#27824C")
```

### Your Turn - Bayes Theorem

What is the probability an article is fake given it has en exclamation point = P(B\|A)

P(B)

```{r}

```

```{r}
# prior_probability_fake <- 
```

P(A\|B)

```{r}

```

```{r}
# probability_has_excl_given_fake <- 
```

P(A)

```{r}

```

```{r}
#probability_has_excl <- 
```

P(B\|A)

```{r}

```

```{r}
news_data |> 
  tabyl(type, title_has_excl) |> 
  adorn_percentages("col")
```

```{r}
news_data %>%
  count(title_has_excl, type) %>%
  group_by(title_has_excl) %>%
  mutate(percent = n / sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x = title_has_excl, y = percent, fill = type)) +
  geom_col(position = "fill") +
  geom_text(aes(label = scales::percent(percent, accuracy = 1)),
            position = position_fill(vjust = 0.5),
            color = "white", fontface = "bold") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c(fake = "#C0392B", real = "#27824C")) +
  labs(x = "Title has exclamation point?", y = "Percentage",
       fill = "Article type",
       title = "Fake vs. real articles, by exclamation point in title") +
  theme_minimal()
```

### Naive Bayes Model

```{r}
model1 <- naiveBayes(type ~ title_has_excl, news_data)
```

```{r}
model1
```

```{r}
news_data <- news_data %>%
  mutate(predicted_type = predict(model1, newdata = .))

news_data %>%
  tabyl(type, predicted_type) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

- **Sensitivity (recall for fake)**: 16/60 ≈ 26.7% — genuinely poor

- **Specificity (recall for real)**: 88/90 ≈ 97.8% — excellent

#### Adding more features (model2)

```{r}
model2 <- naiveBayes(type ~ title_has_excl + title_words, news_data)
```

```{r}
model2
```

```{r}
news_data <- news_data %>%
  mutate(predicted_type = predict(model2, newdata = .))

news_data %>%
  tabyl(type, predicted_type) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

```{r}
ggplot(news_data, aes(x = title_words)) +
  geom_density()
```

```{r}

plot_normal(mean = 12.31667, sd = 3.743884) +
  stat_function(fun = dnorm, args = list(mean = 10.42222, sd = 3.204554),
                color = "forestgreen", linewidth = 1) +
  annotate("text", x = 12.3, y = dnorm(12.3, 12.31667, 3.743884) + 0.01,
           label = "fake", color = "black") +
  annotate("text", x = 10.4, y = dnorm(10.4, 10.42222, 3.204554) + 0.01,
           label = "real", color = "forestgreen") +
  labs(title = "title_words by article type (naive Bayes' assumed Normal curves)",
       x = "title_words", y = "density")
```

#### Adding more features (model3)

```{r}
model3 <- naiveBayes(type ~ title_has_excl + title_words + title_caps_percent + negative,
                     data = news_data)
model3
```

```{r}
news_data <- news_data %>%
  mutate(predicted_type = predict(model3, newdata = .))

news_data %>%
  tabyl(type, predicted_type) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

| Model | Sensitivity (catching fake) | Specificity (catching real) | Accuracy |
|------------------|------------------|------------------|------------------|
| Model 1 (`title_has_excl`) | 26.67% | 97.78% | 69.3% |
| Model 2 (+ `title_words`) | 36.67% | 90.00% | 68.7% |
| Model 3 (+ `title_caps_percent`, `negative`) | **48.33%** | 83.33% | **69.33%** |

#### Adding more features (model4)

```{r}
model4_data <- news_data %>%
  select(-any_of(c("title", "text", "url", "authors", "id", "predicted_type")))

model4 <- naiveBayes(type ~ ., data = model4_data)
```

```{r}
news_data <- news_data %>%
  mutate(predicted_type = predict(model4, newdata = .))

news_data %>%
  tabyl(type, predicted_type) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

#### Your turn - adding more features

Create a model with a custom feature list and test for accuracy with a confusion matrix

```{r}

```

```{r}

```

#### Strategic features

```{r}
candidate_features <- c("title_has_excl", "title_words", "title_caps_percent",
                        "title_excl_percent", "anger", "negative",
                        "text_syllables_per_word", "fear")

evaluate_formula <- function(features, data, k = 5, seed = 84735) {
  set.seed(seed)
  folds <- sample(rep(1:k, length.out = nrow(data)))
  formula <- as.formula(paste("type ~", paste(features, collapse = " + ")))
  
  fold_results <- map_dfr(1:k, function(fold) {
    train <- data[folds != fold, ]
    test  <- data[folds == fold, ]
    m <- naiveBayes(formula, data = train)
    preds <- predict(m, newdata = test)
    cm <- table(factor(test$type, levels = c("fake","real")),
                factor(preds, levels = c("fake","real")))
    tibble(
      sensitivity = cm["fake","fake"] / sum(cm["fake",]),
      specificity = cm["real","real"] / sum(cm["real",]),
      accuracy    = sum(diag(cm)) / sum(cm)
    )
  })
  
  fold_results %>%
    summarise(across(everything(), mean, na.rm = TRUE)) %>%
    mutate(youdens_j = sensitivity + specificity - 1,
           n_features = length(features),
           features = paste(features, collapse = ", "))
}

all_subsets <- unlist(
  lapply(seq_along(candidate_features), function(k) combn(candidate_features, k, simplify = FALSE)),
  recursive = FALSE
)

subset_results <- map_dfr(all_subsets, ~evaluate_formula(.x, model4_data))

subset_results %>% arrange(desc(youdens_j)) %>% head(10)

subset_results %>%
  mutate(weighted_score = 0.7 * sensitivity + 0.3 * specificity) %>%
  arrange(desc(weighted_score)) %>%
  head(10)
```

#### Your turn - Create a visual from subset_results
