---
title: "naive bayes penguins"
format: html
editor: visual
---

### Libraries

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(bayesrules)
library(janitor)
library(e1071)
options(scipen = 99)
```

### Story

We’ll start our naive Bayes classification with just a single penguin. Suppose an Antarctic researcher comes across a penguin that weighs less than 4200g with a 195mm-long flipper and 50mm-long bill. Our goal is to help this researcher identify the species of this penguin, Adelie, Chinstrap, or Gentoo.

#### Let's Plot!

Make a visualization that will help the researcher determine the species based on the characteristics described above

```{r}
penguins <- bayesrules::penguins_bayes

penguins <- penguins %>%
  mutate(above_average_weight = if_else(above_average_weight == 1, "yes", "no"))
```

```{r}

```

```{r}

```

```{r}

```

Our original Bayesian ways

P(B\|A)

- P(Chinstrap \| below average weight)

```{r}
penguins %>% 
  tabyl(above_average_weight, species) %>% 
  adorn_percentages("row")
```

#### Naive Bayes Classification!

```{r}
naive_model_hints <- naiveBayes(species ~ flipper_length_mm + bill_length_mm + above_average_weight, data = penguins)
```

Our penguin

```{r}
our_penguin <- data.frame(bill_length_mm = 50, flipper_length_mm = 195, above_average_weight = "no")
```

Ask our model to make a prediction of what species our penguin is

```{r}
predict(naive_model_hints, newdata = our_penguin, type = "raw")
```

"There's approximately a 99.94% probability this penguin is a Chinstrap, given its flipper length, bill length, and weight category"

Test our model fro accuracy with confusion matrix

```{r}
penguins <- penguins %>% 
  mutate(predicted_species = predict(naive_model_hints, newdata = .))
```

```{r}
penguins %>% 
  tabyl(species, predicted_species) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

```{r}
# Flag whether each prediction matched the true species
penguins <- penguins %>%
  filter(!is.na(species), !is.na(predicted_species)) %>%
  mutate(correct = species == predicted_species)

# Scatterplot: true species by color, misclassified points marked with an X
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm,
                     color = species, shape = correct)) +
  geom_point(size = 2, alpha = 0.7) +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 4)) +
  labs(title = "Misclassified penguins (X) vs. correctly classified (circle)",
       shape = "Correctly classified?", color = "True species") +
  theme_minimal()
```

```{r}
naive_model_hints
```

## Part 2 - Under the hood of Naive Bayes

### Adelie

#### Species

```{r}
prior_prob_adelie <- penguins %>% 
  tabyl(species) %>% 
  filter(species == "Adelie") %>% 
  pull(percent)

prior_prob_adelie
```

#### Flipper Length - Probability density function

- first argument = our penguins flipper length

- second argument = mean flipper length for adelie penguins

- third argument = standard deviation in flipper length for adelie penguins

```{r}
adelie_flipper_mean <- naive_model_hints$tables$flipper_length_mm["Adelie", 1]
adelie_flipper_mean
adelie_flipper_sd   <- naive_model_hints$tables$flipper_length_mm["Adelie", 2]
adelie_flipper_sd
density_flipper_given_adelie <- dnorm(195, mean = adelie_flipper_mean, sd = adelie_flipper_sd)
density_flipper_given_adelie 
```

Given the penguin were an Adelie, a flipper length of 195mm would have a density of about 0.0455 under Adelie's flipper-length distribution

- "195mm is a fairly typical flipper length for an Adelie — the Adelie flipper-length distribution has a density of 0.045296 there, meaning it's reasonably close to Adelie's average of 190mm."

What happened here

```{r}
adelie <- penguins %>% filter(species == "Adelie")

flipper_stats <- adelie %>%
  summarise(mean = mean(flipper_length_mm, na.rm = TRUE),
            sd   = sd(flipper_length_mm, na.rm = TRUE))

highlight_value   <- 195
highlight_density <- dnorm(highlight_value, flipper_stats$mean, flipper_stats$sd)

ggplot(adelie, aes(x = flipper_length_mm)) +
  geom_density(color = "blue", linewidth = 1) +
  annotate("point", x = highlight_value, y = highlight_density, color = "red", size = 3) +
  geom_vline(xintercept = highlight_value, linetype = "dashed", color = "red") +
  geom_vline(xintercept = flipper_stats$mean + c(-1, 1) * flipper_stats$sd, linetype = "dotted") +
  annotate("text",
           x = flipper_stats$mean + c(-1, 1) * flipper_stats$sd + c(-1, 1),
           y = 0.01, label = c("-1 SD", "+1 SD"), hjust = c(1, 0)) +
  annotate("text", x = highlight_value + 2, y = highlight_density,
           label = paste0("195 mm\nDensity: ", round(highlight_density, 5)),
           hjust = 0, color = "red") +
  labs(title = "Normal Distribution of Flipper Lengths for Adelie Penguins",
       subtitle = paste("Mean =", round(flipper_stats$mean, 2), "mm, SD =", round(flipper_stats$sd, 2), "mm"),
       x = "Flipper Length (mm)", y = "Density") +
  theme_minimal()
```

```{r}
ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram()
```

#### Bill Length

```{r}
adelie_bill_mean <- naive_model_hints$tables$bill_length_mm["Adelie", 1]
adelie_bill_mean
adelie_bill_sd   <- naive_model_hints$tables$bill_length_mm["Adelie", 2]
adelie_bill_sd
density_bill_given_adelie <- dnorm(50, mean = adelie_bill_mean, sd = adelie_bill_sd)
density_bill_given_adelie 
```

#### Above Average Weight

```{r}
prob_under_average_weight_given_adelie <- penguins %>%
  filter(species == "Adelie") %>%
  count(above_average_weight) %>%
  mutate(prop = n / sum(n)) %>%
  filter(above_average_weight == "no") %>%
  pull(prop)

prob_under_average_weight_given_adelie
```

P(Adelie) \* P(195 mm flipper \| Adelie ) \* P(50 mm bill \| Adelie ) \* P(under average weight \| Adelie)

```{r}
probs_adelie <- prior_prob_adelie * density_flipper_given_adelie * density_bill_given_adelie * prob_under_average_weight_given_adelie
probs_adelie

```

### Chinstrap

#### Species

```{r}
prior_prob_chinstrap <- penguins %>% 
  tabyl(species) %>% 
  filter(species == "Chinstrap") %>% 
  pull(percent)

prior_prob_chinstrap
```

#### Flipper Length

```{r}
chinstrap_flipper_mean <- naive_model_hints$tables$flipper_length_mm["Chinstrap", 1]
chinstrap_flipper_mean
chinstrap_flipper_sd   <- naive_model_hints$tables$flipper_length_mm["Chinstrap", 2]
chinstrap_flipper_sd
density_flipper_given_chinstrap <- dnorm(195, mean = chinstrap_flipper_mean, sd = chinstrap_flipper_sd)
density_flipper_given_chinstrap
```

#### Bill Length

```{r}
chinstrap_bill_mean <- naive_model_hints$tables$bill_length_mm["Chinstrap", 1]
chinstrap_bill_mean
chinstrap_bill_sd   <- naive_model_hints$tables$bill_length_mm["Chinstrap", 2]
chinstrap_bill_sd
density_bill_given_chinstrap <- dnorm(50, mean = chinstrap_bill_mean, sd = chinstrap_bill_sd)
density_bill_given_chinstrap 
```

#### Above Average Weight

```{r}
prob_under_average_weight_given_chinstrap  <- penguins %>%
  filter(species == "Chinstrap") %>%
  count(above_average_weight) %>%
  mutate(prop = n / sum(n)) %>%
  filter(above_average_weight == "no") %>%
  pull(prop)

prob_under_average_weight_given_chinstrap 
```

P(Chinstrap) \* P(195 mm flipper \| Chinstrap ) \* P(50 mm bill \| Chinstrap ) \* P(under average weight \| Chinstrap)

```{r}
probs_chinstrap <- prior_prob_chinstrap * density_flipper_given_chinstrap * density_bill_given_chinstrap * prob_under_average_weight_given_chinstrap
probs_chinstrap
```

### Your turn - Find Gentoo

#### Species

```{r}

```

#### Flipper Length

```{r}

```

#### Bill Length

```{r}

```

#### Above Average Weight

```{r}

```

P(Gentoo) \* P(195 mm flipper \| Gentoo ) \* P(50 mm bill \| Gentoo ) \* P(under average weight \| Gentoo)

```{r}

```

### Final

```{r}
# total_prob <- probs_adelie + probs_chinstrap + probs_gentoo
# 
# posterior_probs <- c(
#   Adelie    = probs_adelie / total_prob,
#   Chinstrap = probs_chinstrap / total_prob,
#   Gentoo    = probs_gentoo / total_prob
# )
# 
# posterior_probs
```

```{r}
#predict(naive_model_hints, newdata = our_penguin, type = "raw")
```

### What makes Naive bayes naive?

```{r}
penguins %>%
  filter(!is.na(bill_length_mm), !is.na(flipper_length_mm)) %>%
  group_by(species) %>%
  summarise(correlation = cor(bill_length_mm, flipper_length_mm))
```

```{r}
library(tidyverse)

# Build both curves on the SAME x-grid per species/feature, so the gap
# between "actual" and "assumed Normal" can be shaded directly rather
# than just visually estimated from two independently-drawn curves.
build_comparison <- function(values, species_name, feature_name, n_grid = 200) {
  values <- values[!is.na(values)]
  x_seq  <- seq(min(values), max(values), length.out = n_grid)
  
  actual_dens  <- density(values, n = n_grid, from = min(x_seq), to = max(x_seq))$y
  assumed_dens <- dnorm(x_seq, mean(values), sd(values))
  
  tibble(value = x_seq, actual = actual_dens, assumed = assumed_dens,
         species = species_name, feature = feature_name)
}

raw_data <- penguins %>%
  filter(!is.na(species)) %>%
  dplyr::select(species, bill_length_mm, flipper_length_mm) %>%
  pivot_longer(cols = c(bill_length_mm, flipper_length_mm),
               names_to = "feature", values_to = "value") %>%
  filter(!is.na(value))

comparison_data <- raw_data %>%
  group_by(species, feature) %>%
  summarise(data = list(build_comparison(value, unique(species), unique(feature))), .groups = "drop") %>%
  pull(data) %>%
  bind_rows()

feature_labels <- c(bill_length_mm = "Bill Length (mm)", flipper_length_mm = "Flipper Length (mm)")

ggplot(comparison_data, aes(x = value)) +
  geom_ribbon(aes(ymin = pmin(actual, assumed), ymax = pmax(actual, assumed)),
              fill = "grey55", alpha = 0.35) +
  geom_line(aes(y = actual, color = "Actual data"), linewidth = 1) +
  geom_line(aes(y = assumed, color = "Naive Bayes' assumed Normal"),
            linewidth = 1, linetype = "dashed") +
  geom_rug(data = raw_data, aes(x = value), sides = "b", alpha = 0.3, inherit.aes = FALSE) +
  facet_grid(feature ~ species, scales = "free",
             labeller = labeller(feature = feature_labels)) +
  scale_color_manual(values = c("Actual data" = "#4C72B0",
                                "Naive Bayes' assumed Normal" = "#DD8452"),
                     name = NULL) +
  labs(title = "How well does naive Bayes' Normal assumption fit the real data?",
       subtitle = "Shaded grey = the gap between what's real (blue) and what the model assumes (orange, dashed)",
       x = NULL, y = "Density") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top", strip.text = element_text(face = "bold"))
```
