---
title: "Rain in Perth"
format: html
editor: visual
---

Welcome to week 8! Today, we are focused on modeling rain based on humidity at 9 am

### Packages

```{r}
library(tidyverse)
library(bayesrules)
library(e1071)
library(janitor)
```

### Data

```{r}
weather <- bayesrules::weather_perth
weather <- weather %>% 
  select(day_of_year, raintomorrow, humidity9am, humidity3pm, raintoday)
```

### EDA

Take 15 minutes to utilize visualizations to consider the relationship between rain tomorrow and the humidity at 9 am today

```{r}

```

```{r}

```

```{r}
weather %>%
  mutate(rain_num = as.numeric(raintomorrow == "Yes")) %>%
  ggplot(aes(x = humidity9am, y = rain_num)) +
  geom_jitter(height = 0.05, alpha = 0.3) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  labs(y = "Rain tomorrow (1 = yes)")
```

### Fit a naive bayes classification model

Only use humidity at 9 am as a predictive feature and evaluate performance. Call it nb_class_model

```{r}

```

```{r}
#nb_class_model
```

```{r}

```

Fit a naive bayes classification model with all the available data as features and evaluate performance. Call it nb_class_model_all

```{r}

```

```{r}

```

```{r}

```

```{r}

```

### What does this model tell us and not tell us?

- 

### Logistic Regression

#### Frequentist approach

```{r}
rain_glm <- glm(raintomorrow ~ humidity9am, data = weather, family = binomial)
summary(rain_glm)
exp(coef(rain_glm))      # odds multipliers
exp(confint(rain_glm))   # 95% confidence intervals on the odds scale
```

**The Wald z-test in `summary()`.** z = 8.95 (0.0476 / 0.0053) with p \< 2e-16, testing H₀: β₁ = 0. Very strong evidence that humidity is associated with rain tomorrow.

**Humidity (odds multiplier 1.049):** each additional percentage point of 9 a.m. humidity multiplies the odds of rain tomorrow by about 1.049, an increase of roughly 4.9%.

A multiplier of 1.049 means the new odds are 1.049 times the old odds, or 104.9% of them. The increase is the part above 100%:

1.0487 - 1 = 0.0487 -\> about 4.9%

**95% confidence interval for humidity (1.038 to 1.060):** this procedure would capture the true odds multiplier 95% of the time across repeated samples. Taken at face value, each humidity point raises the odds of rain by somewhere between about 3.8% and 6.0%. The whole interval sits above 1, which means no effect, so humidity is clearly associated with higher odds of rain. This is a statement about the long-run reliability of the method, not a 95% probability that this particular interval contains the true value. The Bayesian credible interval makes that claim.

#### Bayesian Approach

```{r}
rain_model <- stan_glm(raintomorrow ~ humidity9am,
                             data = weather, family = binomial,
                             prior_intercept = normal(-1.4, 0.7),
                             prior = normal(0.07, 0.035),
                             chains = 4, iter = 5000*2, seed = 84735)
```

```{r}
summary(rain_model, digits = 4)
```

### App

```{r}
## ============================================================
## Bayesian Logistic Regression Explorer: Will it rain in Perth?
## Follows Bayes Rules! Chapter 13 (Logistic Regression)
##
## Tab 1: Odds, log-odds & probability -- the three "scales" the
##        whole chapter moves between (Section 13.1 / Figure 13.1).
## Tab 2: Tune the prior (beta0c, beta1) -- fully live, no MCMC,
##        just simulating directly from the priors (Figure 13.2).
## Tab 3: Fit the REAL posterior via stan_glm(family = binomial) on
##        the weather_perth data -- real MCMC, so it's behind a button.
## Tab 4: Predict & classify -- reuses the fitted posterior draws, so
##        it's fully live: pick a humidity, pick a classification
##        cutoff, and watch sensitivity/specificity trade off.
##
## Run with: shiny::runApp("app.R")
## ============================================================

library(shiny)
library(tidyverse)
library(bayesrules)   # weather_perth data + plot_normal()
library(rstanarm)     # stan_glm() for the real MCMC fit

## ------------------------------------------------------------
## Data (loaded once; never changes while the app runs)
## ------------------------------------------------------------

weather <- bayesrules::weather_perth %>%
  select(day_of_year, raintomorrow, humidity9am, humidity3pm, raintoday) %>%
  mutate(y = as.numeric(raintomorrow == "Yes"))

mean_h     <- mean(weather$humidity9am)
h_range    <- range(weather$humidity9am)
base_rate  <- mean(weather$y)
n_days     <- nrow(weather)

# "Data alone" reference for beta1: an ordinary maximum-likelihood
# logistic fit, ignoring the prior entirely. Its estimate +/- standard
# error is the frequentist stand-in for "what the data alone says".
mle_model     <- glm(raintomorrow ~ humidity9am, data = weather, family = binomial)
mle_coef      <- summary(mle_model)$coefficients
mle_beta1_est <- unname(mle_coef["humidity9am", "Estimate"])
mle_beta1_se  <- unname(mle_coef["humidity9am", "Std. Error"])

# Observed rain rate within humidity brackets (the book's Figure 13.3 idea)
binned_rain <- weather %>%
  mutate(bracket = cut(humidity9am, breaks = seq(0, 100, by = 10), include.lowest = TRUE)) %>%
  group_by(bracket) %>%
  summarise(humidity = mean(humidity9am), rain_rate = mean(y), n = n(), .groups = "drop") %>%
  filter(n >= 5)

app_css <- tags$head(tags$style(HTML("
  .story-panel {
    background: #eef4fb; border-left: 4px solid #4C72B0;
    border-radius: 4px; padding: 10px 16px; margin-bottom: 14px;
  }
  .interp-panel {
    background: #f2f8f0; border-left: 4px solid #55A868;
    border-radius: 4px; padding: 10px 16px; margin-top: 10px;
  }
  .concept-panel {
    background: #f7f2fb; border-left: 4px solid #8B5CF6;
    border-radius: 4px; padding: 10px 16px; margin-top: 10px;
    margin-bottom: 10px; font-size: 0.95em;
  }
  .code-box {
    background: #2b2b2b; color: #eaeaea; font-family: monospace;
    font-size: 13px; padding: 12px 16px; border-radius: 6px;
    white-space: pre-wrap; margin-top: 8px;
  }
  .verdict-rain { background:#e8f0fb; border-left:4px solid #2c6fbb; }
  .verdict-dry  { background:#fdf3e3; border-left:4px solid #d9962b; }
  .verdict { border-radius:6px; padding:14px; margin-top:10px; font-size:1.1em; }
")))

## ------------------------------------------------------------
## Helpers
## ------------------------------------------------------------

stan_call_text <- function(m0, s0, m1, s1, prior_pd) {
  sprintf(
    paste0(
      "stan_glm(\n",
      "  raintomorrow ~ humidity9am, data = weather,\n",
      "  family = binomial,\n",
      "  prior_intercept = normal(%s, %s),\n",
      "  prior = normal(%s, %s),\n",
      "  chains = 4, iter = 2000, seed = 84735%s)"
    ),
    m0, s0, m1, s1,
    if (prior_pd) ",\n  prior_PD = TRUE" else ""
  )
}

pct <- function(x, digits = 0) sprintf(paste0("%.", digits, "f%%"), 100 * x)

# Classification metrics for a given probability vector, truth vector, cutoff
class_metrics <- function(prob, truth, cutoff) {
  pred <- as.numeric(prob >= cutoff)
  a <- sum(truth == 0 & pred == 0); b <- sum(truth == 0 & pred == 1)
  c <- sum(truth == 1 & pred == 0); d <- sum(truth == 1 & pred == 1)
  list(
    cm = matrix(c(a, c, b, d), nrow = 2,
                dimnames = list(`Actual` = c("No rain", "Rain"), `Predicted` = c("No rain", "Rain"))),
    sensitivity = if ((c + d) > 0) d / (c + d) else NA_real_,
    specificity = if ((a + b) > 0) a / (a + b) else NA_real_,
    accuracy    = (a + d) / (a + b + c + d)
  )
}

## ------------------------------------------------------------
## UI
## ------------------------------------------------------------

ui <- navbarPage(
  title = "Bayesian Logistic Regression: Rain in Perth",
  header = app_css,
  
  ## ---------------- TAB 1: Odds primer ----------------
  tabPanel(
    "1. Odds & Probability",
    fluidPage(
      div(class = "story-panel",
          strong("Why start here:"), " logistic regression models a yes/no ",
          "outcome, but it never draws a straight line through probabilities ",
          "-- a straight line would eventually predict probabilities below 0 ",
          "or above 1. Instead it draws a straight line on the ",
          em("log-odds"), " scale, then converts back. This tab lets you ",
          "feel those conversions before any data enters."
      ),
      h3("Converting between probability, odds, and log-odds"),
      sliderInput("prob_in", "Probability of rain tomorrow (\u03c0)",
                  min = 0.01, max = 0.99, value = 0.5, step = 0.01, width = "500px"),
      uiOutput("oddsStatement"),
      plotOutput("oddsCurve", height = "250px"),
      div(class = "concept-panel",
          strong("The three scales, side by side:"), " probability lives between ",
          "0 and 1; odds (\u03c0 / (1 - \u03c0)) live between 0 and infinity; ",
          "log-odds live on the whole number line, negative or positive. ",
          "That last fact is the whole point: a straight line \u03b2\u2080 + \u03b2\u2081X ",
          "can take any real value, so it can only sensibly model something ",
          "that also lives on the whole number line -- log-odds, not ",
          "probability."
      ),
      hr(),
      h3("One straight line, three different pictures (the book's Figure 13.1)"),
      fluidRow(
        column(6, sliderInput("demo_b0", "\u03b2\u2080 (log-odds at 0% humidity)", -8, 2, -4, step = 0.1)),
        column(6, sliderInput("demo_b1", "\u03b2\u2081 (change in log-odds per humidity point)", 0, 0.2, 0.1, step = 0.005))
      ),
      uiOutput("demoStatement"),
      fluidRow(
        column(4, plotOutput("demoLogOdds", height = "230px")),
        column(4, plotOutput("demoOdds", height = "230px")),
        column(4, plotOutput("demoProb", height = "230px"))
      )
    )
  ),
  
  ## ---------------- TAB 2: Tune the prior ----------------
  tabPanel(
    "2. Tune the Prior",
    sidebarLayout(
      sidebarPanel(
        div(class = "story-panel",
            strong("The scenario:"), " we're in Perth, Australia, trying to ",
            "predict whether it rains tomorrow from today's 9 a.m. humidity. ",
            "Vaguely, we believe rain is fairly uncommon on an average day ",
            "(maybe 20%), and more likely after a humid day -- but we're ",
            "foggy on how much more."
        ),
        h4("beta0c ~ Normal(m0, s0), on the log-odds scale"),
        sliderInput("m0", "m0 (log-odds of rain on an average-humidity day)", -4, 1, -1.4, step = 0.1),
        sliderInput("s0", "s0 (uncertainty around that)", 0.1, 2, 0.7, step = 0.05),
        h4("beta1 ~ Normal(m1, s1)"),
        sliderInput("m1", "m1 (typical change in log-odds per humidity point)", -0.05, 0.2, 0.07, step = 0.005),
        sliderInput("s1", "s1 (uncertainty around that)", 0.005, 0.1, 0.035, step = 0.005),
        actionButton("reset_prior", "Reset to the text's priors"),
        div(class = "concept-panel",
            strong("Priors live on the log-odds scale:"), " that's where \u03b2\u2080 and ",
            "\u03b2\u2081 are allowed to be any real number, so Normal priors are the ",
            "natural choice. The statements on the right translate your ",
            "log-odds choices into the odds and probability scales, which ",
            "are far easier to sanity-check."
        )
      ),
      mainPanel(
        uiOutput("priorStatement"),
        fluidRow(
          column(4, plotOutput("beta0cPlot", height = "210px")),
          column(4, plotOutput("beta1Plot", height = "210px")),
          column(4, plotOutput("impliedProbPlot", height = "210px"))
        ),
        h4("100 prior plausible S-curves: probability of rain vs. 9 a.m. humidity"),
        plotOutput("priorCurvesPlot", height = "280px"),
        div(class = "concept-panel",
            strong("Prior predictive check:"), " every S-curve above is one ",
            "version of the world your priors consider plausible. The ",
            "histogram below asks a different but equally important ",
            "question: if each of those worlds produced a full year of ",
            "weather, what share of days would be rainy? If that range ",
            "looks absurd for Perth (a dry-summer city), retune."
        ),
        plotOutput("priorRainRatePlot", height = "230px"),
        h4("Equivalent stan_glm() code (prior only, prior_PD = TRUE)"),
        div(class = "code-box", textOutput("priorCode"))
      )
    )
  ),
  
  ## ---------------- TAB 3: Fit the posterior ----------------
  tabPanel(
    "3. Fit the Posterior",
    sidebarLayout(
      sidebarPanel(
        div(class = "story-panel",
            strong("This runs real MCMC"), " on ", n_days, " days of Perth ",
            "weather, using the priors you tuned on Tab 2. Expect a few ",
            "seconds -- Stan is genuinely sampling the posterior."
        ),
        p(strong("Current priors (from Tab 2):")),
        verbatimTextOutput("currentPriorSummary"),
        actionButton("fit_posterior", "Fit the posterior model", class = "btn-primary"),
        hr(),
        sliderInput("ci_level", "Credible interval level (%)", 50, 99, 80, step = 1),
        helpText("4 chains x 2000 iterations (fewer than the book's 10,000 ",
                 "to keep this responsive in a live demo).")
      ),
      mainPanel(
        uiOutput("posteriorStatus"),
        conditionalPanel(
          condition = "output.modelFitted == true",
          h4("What the data says about humidity"),
          uiOutput("posteriorStatement"),
          tableOutput("posteriorTable"),
          div(class = "concept-panel",
              strong("Three scales again:"), " the table reports beta1 on the ",
              "log-odds scale (a plain slope) and on the odds scale ",
              "(e^beta1, a multiplier). A multiplier of 1.05 means each extra ",
              "humidity point raises the odds of rain by 5%. The probability ",
              "scale is skipped on purpose -- because the S-curve is ",
              "nonlinear, there's no single 'per point' change in probability."
          ),
          h4("Prior belief vs. data alone vs. posterior, for beta1"),
          plotOutput("beta1ComparePlot", height = "260px"),
          div(class = "concept-panel",
              strong("Reading the three curves:"), " blue = your prior belief ",
              "about beta1; orange = what an ordinary (non-Bayesian) logistic ",
              "fit says from the data alone; green = the posterior. With ",
              n_days, " real days, the data-alone curve is narrow, so the ",
              "posterior lands close to it unless your prior is very tightly ",
              "specified -- the same weight-of-evidence story as every ",
              "model this semester."
          ),
          h4("100 posterior plausible S-curves vs. the observed rain rate"),
          plotOutput("posteriorCurvesPlot", height = "280px"),
          div(class = "concept-panel",
              strong("The points are real data:"), " each dot is the actual ",
              "share of days that were followed by rain within a 10-point ",
              "humidity bracket. A good model's curves should thread through ",
              "them."
          ),
          h4("Equivalent stan_glm() code (the exact call used)"),
          div(class = "code-box", textOutput("posteriorCode")),
          h4("MCMC diagnostics"),
          verbatimTextOutput("diagnostics"),
          div(class = "concept-panel",
              strong("Rhat and n_eff:"), " Rhat compares your chains against ",
              "each other -- near 1.00 means they converged on the same ",
              "posterior; values noticeably above 1 (worry past ~1.01-1.05) ",
              "mean the simulation hasn't settled. n_eff is roughly how many ",
              "genuinely independent draws you got. Both are about whether ",
              "the ", em("simulation"), " worked, not whether you have enough ",
              em("data"), "."
          )
        )
      )
    )
  ),
  
  ## ---------------- TAB 4: Predict & classify ----------------
  tabPanel(
    "4. Predict & Classify",
    sidebarLayout(
      sidebarPanel(
        div(class = "story-panel",
            strong("Requires a fitted model from Tab 3."), " Once fitted, ",
            "this tab is fully live -- no new MCMC, just reusing the ",
            "posterior draws already in memory."
        ),
        sliderInput("h_predict", "Today's 9 a.m. humidity (%)", 10, 100, 99, step = 1),
        sliderInput("cutoff", "Classification cutoff c", 0.05, 0.95, 0.5, step = 0.05),
        div(class = "concept-panel",
            strong("What's the cutoff?"), " the model gives a probability of ",
            "rain; a classification needs a yes/no. The cutoff is the ",
            "probability above which you say \"rain.\" 0.5 is the natural ",
            "default, but nothing forces it -- if getting caught in the rain ",
            "is worse than carrying a useless umbrella, lower it."
        )
      ),
      mainPanel(
        uiOutput("predictionStatus"),
        conditionalPanel(
          condition = "output.modelFitted == true",
          h4("Predicting tomorrow, given today's humidity"),
          uiOutput("verdict"),
          plotOutput("predictionPlot", height = "230px"),
          div(class = "code-box", textOutput("predictionCode")),
          hr(),
          h4("How good is this at classifying all 1,000 days at once?"),
          uiOutput("metricsStatement"),
          tableOutput("confusionTable"),
          plotOutput("tradeoffPlot", height = "280px"),
          div(class = "concept-panel",
              strong("The trade-off:"), " lowering the cutoff makes it easier ",
              "to say \"rain,\" which catches more of the real rainy days ",
              "(sensitivity up) but also cries wolf on more dry days ",
              "(specificity down). Raising it does the opposite. There is no ",
              "setting that wins on both -- the right one depends on which ",
              "mistake costs you more."
          ),
          div(class = "code-box", textOutput("classificationCode"))
        )
      )
    )
  )
)

## ------------------------------------------------------------
## Server
## ------------------------------------------------------------

server <- function(input, output, session) {
  
  ## ===== TAB 1: Odds primer =====
  
  output$oddsStatement <- renderUI({
    p <- input$prob_in
    odds <- p / (1 - p)
    lo <- log(odds)
    verbal <- if (abs(odds - 1) < 0.005) {
      "rain and no rain are exactly equally likely"
    } else if (odds > 1) {
      sprintf("rain is %.2f times as likely as no rain", odds)
    } else {
      sprintf("rain is only %.2f times as likely as no rain (no rain is %.2f times as likely as rain)",
              odds, 1 / odds)
    }
    div(class = "interp-panel",
        sprintf("A probability of %s corresponds to odds of %.3f, so %s. On the log-odds scale that's %.3f.",
                pct(p), odds, verbal, lo)
    )
  })
  
  output$oddsCurve <- renderPlot({
    p <- input$prob_in
    lo <- log(p / (1 - p))
    x <- seq(-6, 6, length.out = 300)
    ggplot(tibble(lo = x, p = plogis(x)), aes(lo, p)) +
      geom_line(color = "#4C72B0", linewidth = 1.1) +
      annotate("point", x = lo, y = p, color = "#C0392B", size = 4) +
      geom_hline(yintercept = 0.5, linetype = "dotted") +
      geom_vline(xintercept = 0, linetype = "dotted") +
      labs(x = "log-odds", y = "probability",
           title = "The logistic S-curve: log-odds in, probability out") +
      theme_minimal(base_size = 13)
  })
  
  demo_df <- reactive({
    x <- seq(0, 100, length.out = 200)
    lo <- input$demo_b0 + input$demo_b1 * x
    tibble(humidity = x, log_odds = lo, odds = exp(lo), prob = plogis(lo))
  })
  
  output$demoStatement <- renderUI({
    b0 <- input$demo_b0; b1 <- input$demo_b1
    tip <- if (b1 > 0 && -b0 / b1 >= 0 && -b0 / b1 <= 100) {
      sprintf("The probability of rain crosses 50%% (odds = 1) at about %.0f%% humidity.", -b0 / b1)
    } else {
      "Within 0-100% humidity this curve never reaches a 50% chance of rain."
    }
    div(class = "interp-panel",
        sprintf(
          paste0(
            "At 0%% humidity the log-odds of rain are %.1f -- odds of %.4f, a %s chance. ",
            "Each extra humidity point adds %.3f to the log-odds, which multiplies the ",
            "odds by e^%.3f = %.3f (a %.1f%% increase). %s"
          ),
          b0, exp(b0), pct(plogis(b0), 1), b1, b1, exp(b1), 100 * (exp(b1) - 1), tip
        )
    )
  })
  
  output$demoLogOdds <- renderPlot({
    ggplot(demo_df(), aes(humidity, log_odds)) +
      geom_line(color = "#4C72B0", linewidth = 1.1) +
      labs(x = "humidity", y = "log(odds of rain)", title = "Log-odds: a straight line") +
      theme_minimal(base_size = 12)
  })
  output$demoOdds <- renderPlot({
    ggplot(demo_df(), aes(humidity, odds)) +
      geom_line(color = "#DD8452", linewidth = 1.1) +
      labs(x = "humidity", y = "odds of rain", title = "Odds: curves upward") +
      theme_minimal(base_size = 12)
  })
  output$demoProb <- renderPlot({
    ggplot(demo_df(), aes(humidity, prob)) +
      geom_line(color = "#55A868", linewidth = 1.1) +
      ylim(0, 1) +
      labs(x = "humidity", y = "probability of rain", title = "Probability: an S-curve") +
      theme_minimal(base_size = 12)
  })
  
  ## ===== TAB 2: Prior tuning (instant) =====
  
  observeEvent(input$reset_prior, {
    updateSliderInput(session, "m0", value = -1.4)
    updateSliderInput(session, "s0", value = 0.7)
    updateSliderInput(session, "m1", value = 0.07)
    updateSliderInput(session, "s1", value = 0.035)
  })
  
  prior_draws <- reactive({
    set.seed(84735)   # fixed seed so curves don't jitter as sliders move
    tibble(beta0c = rnorm(100, input$m0, input$s0),
           beta1  = rnorm(100, input$m1, input$s1),
           draw   = seq_len(100))
  })
  
  output$priorStatement <- renderUI({
    m0 <- input$m0; s0 <- input$s0; m1 <- input$m1; s1 <- input$s1
    lo_lo <- m0 - 1.96 * s0; lo_hi <- m0 + 1.96 * s0
    b_lo  <- m1 - 1.96 * s1; b_hi  <- m1 + 1.96 * s1
    tip <- if (m1 > 0) {
      x50 <- mean_h - m0 / m1
      if (x50 >= 0 && x50 <= 100) sprintf("Your best-guess curve crosses a 50%% chance of rain at about %.0f%% humidity.", x50)
      else "Your best-guess curve never reaches a 50% chance of rain within 0-100% humidity."
    } else "With a zero or negative best-guess slope, your prior doesn't expect humidity to raise the chance of rain."
    
    div(class = "interp-panel",
        sprintf(
          paste0(
            "On an average-humidity day (~%.0f%%), your best guess is a %s chance of rain ",
            "(log-odds %.1f, odds %.2f), and you'd be unsurprised by anything between about %s ",
            "and %s (log-odds %.1f to %.1f). For every one-point rise in humidity you expect the ",
            "log-odds to change by about %.3f -- multiplying the odds by e^%.3f = %.3f (95%% plausible ",
            "range of multipliers: %.3f to %.3f). A 15-point rise multiplies the odds by about %.2f. %s"
          ),
          mean_h, pct(plogis(m0), 0), m0, exp(m0), pct(plogis(lo_lo), 0), pct(plogis(lo_hi), 0),
          lo_lo, lo_hi, m1, m1, exp(m1), exp(b_lo), exp(b_hi), exp(15 * m1), tip
        )
    )
  })
  
  output$beta0cPlot <- renderPlot({
    plot_normal(mean = input$m0, sd = input$s0) +
      labs(x = "beta0c (log-odds, avg. humidity)", y = "pdf") + theme_minimal(base_size = 11)
  })
  output$beta1Plot <- renderPlot({
    plot_normal(mean = input$m1, sd = input$s1) +
      labs(x = "beta1 (log-odds per humidity point)", y = "pdf") + theme_minimal(base_size = 11)
  })
  output$impliedProbPlot <- renderPlot({
    set.seed(1)
    tibble(p = plogis(rnorm(20000, input$m0, input$s0))) %>%
      ggplot(aes(p)) +
      geom_density(fill = "#55A868", alpha = 0.4) +
      xlim(0, 1) +
      labs(x = "implied chance of rain, avg. humidity", y = "density") +
      theme_minimal(base_size = 11)
  })
  
  output$priorCurvesPlot <- renderPlot({
    x_seq <- seq(10, 100, length.out = 60)
    prior_draws() %>%
      crossing(x = x_seq) %>%
      mutate(prob = plogis(beta0c + beta1 * (x - mean_h))) %>%
      ggplot(aes(x, prob, group = draw)) +
      geom_line(alpha = 0.2, color = "#4C72B0") +
      ylim(0, 1) +
      labs(x = "humidity at 9 a.m.", y = "probability of rain tomorrow") +
      theme_minimal(base_size = 13)
  })
  
  output$priorRainRatePlot <- renderPlot({
    set.seed(84735)
    d <- prior_draws()
    rates <- map_dbl(seq_len(nrow(d)), function(i) {
      p <- plogis(d$beta0c[i] + d$beta1[i] * (weather$humidity9am - mean_h))
      mean(rbinom(n_days, 1, p))
    })
    ggplot(tibble(rate = rates), aes(rate)) +
      geom_histogram(color = "white", fill = "#4C72B0", bins = 20) +
      geom_vline(xintercept = base_rate, linetype = "dashed", color = "#C0392B") +
      labs(x = "proportion of rainy days in a simulated dataset", y = "count",
           title = sprintf("Prior-simulated rain rates (red dashed = actual observed rate, %s)", pct(base_rate))) +
      theme_minimal(base_size = 12)
  })
  
  output$priorCode <- renderText({
    stan_call_text(input$m0, input$s0, input$m1, input$s1, prior_pd = TRUE)
  })
  
  ## ===== TAB 3: Real posterior fit =====
  
  output$currentPriorSummary <- renderText({
    sprintf("beta0c ~ Normal(%s, %s)\nbeta1  ~ Normal(%s, %s)", input$m0, input$s0, input$m1, input$s1)
  })
  
  fitted_model <- reactiveVal(NULL)
  fitted_draws <- reactiveVal(NULL)
  
  observeEvent(input$fit_posterior, {
    withProgress(message = "Running MCMC (stan_glm)...", value = 0.3, {
      m <- stan_glm(
        raintomorrow ~ humidity9am, data = weather, family = binomial,
        prior_intercept = normal(input$m0, input$s0),
        prior = normal(input$m1, input$s1),
        chains = 4, iter = 2000, seed = 84735, refresh = 0
      )
      incProgress(0.6)
      fitted_model(m)
      fitted_draws(as.data.frame(m))
    })
  })
  
  output$modelFitted <- reactive({ !is.null(fitted_model()) })
  outputOptions(output, "modelFitted", suspendWhenHidden = FALSE)
  
  output$posteriorStatus <- renderUI({
    if (is.null(fitted_model())) p(em("Click \"Fit the posterior model\" to run MCMC on the real Perth data."))
  })
  
  output$posteriorStatement <- renderUI({
    req(fitted_draws())
    d <- fitted_draws(); lvl <- input$ci_level / 100
    b1 <- d$humidity9am
    ci <- quantile(b1, c((1 - lvl) / 2, 1 - (1 - lvl) / 2))
    med <- median(b1)
    vs_prior <- if (med < input$m1 - 0.005) "smaller than" else if (med > input$m1 + 0.005) "larger than" else "about the same as"
    x50 <- -median(d$`(Intercept)`) / med
    tip <- if (x50 >= 0 && x50 <= 100) sprintf(" The median curve reaches a 50%% chance of rain at roughly %.0f%% humidity.", x50)
    else " Within 0-100% humidity, the median curve never reaches a 50% chance of rain."
    div(class = "interp-panel",
        sprintf(
          paste0(
            "For every one-point increase in today's 9 a.m. humidity, there's a %.0f%% posterior chance the ",
            "log-odds of rain rise by between %.4f and %.4f -- equivalently, the odds of rain rise by between ",
            "%.1f%% and %.1f%%. That median rate (%.4f) is %s your prior mean (%.3f). Every 15-point rise in ",
            "humidity multiplies the odds by roughly %.2f to %.2f.%s"
          ),
          100 * lvl, ci[1], ci[2], 100 * (exp(ci[1]) - 1), 100 * (exp(ci[2]) - 1),
          med, vs_prior, input$m1, exp(15 * ci[1]), exp(15 * ci[2]), tip
        )
    )
  })
  
  output$posteriorTable <- renderTable({
    req(fitted_draws())
    d <- fitted_draws(); lvl <- input$ci_level / 100
    lo_q <- (1 - lvl) / 2; hi_q <- 1 - lo_q
    summarize_param <- function(x, name) {
      tibble(
        parameter = name,
        `log-odds scale (median)` = median(x),
        `lower` = quantile(x, lo_q), `upper` = quantile(x, hi_q),
        `odds scale (median)` = exp(median(x)),
        `odds lower` = exp(quantile(x, lo_q)), `odds upper` = exp(quantile(x, hi_q))
      )
    }
    bind_rows(
      summarize_param(d$`(Intercept)`, "(Intercept)  [raw: log-odds at 0% humidity]"),
      summarize_param(d$humidity9am, "humidity9am")
    )
  }, digits = 4)
  
  output$beta1ComparePlot <- renderPlot({
    req(fitted_draws())
    post <- fitted_draws()$humidity9am
    x_lo <- min(input$m1 - 3 * input$s1, mle_beta1_est - 4 * mle_beta1_se, min(post))
    x_hi <- max(input$m1 + 3 * input$s1, mle_beta1_est + 4 * mle_beta1_se, max(post))
    xs <- seq(x_lo, x_hi, length.out = 500)
    pd <- density(post, n = 500)
    all_df <- bind_rows(
      tibble(x = xs, density = dnorm(xs, input$m1, input$s1), curve = "Prior belief"),
      tibble(x = xs, density = dnorm(xs, mle_beta1_est, mle_beta1_se), curve = "Data alone (MLE)"),
      tibble(x = pd$x, density = pd$y, curve = "Posterior")
    ) %>% mutate(curve = factor(curve, levels = c("Prior belief", "Data alone (MLE)", "Posterior")))
    cols <- c("Prior belief" = "#4C72B0", "Data alone (MLE)" = "#DD8452", "Posterior" = "#55A868")
    ggplot(all_df, aes(x, density, color = curve, fill = curve)) +
      geom_area(alpha = 0.25, position = "identity") +
      geom_line(linewidth = 1.1) +
      scale_color_manual(values = cols) + scale_fill_manual(values = cols) +
      labs(x = "beta1 (change in log-odds of rain per humidity point)", y = "density", color = NULL, fill = NULL) +
      theme_minimal(base_size = 13) + theme(legend.position = "top")
  })
  
  output$posteriorCurvesPlot <- renderPlot({
    req(fitted_draws())
    d <- fitted_draws() %>% slice_sample(n = 100) %>% mutate(draw = row_number())
    x_seq <- seq(10, 100, length.out = 60)
    curves <- d %>% crossing(x = x_seq) %>%
      mutate(prob = plogis(`(Intercept)` + humidity9am * x))
    ggplot() +
      geom_line(data = curves, aes(x, prob, group = draw), alpha = 0.2, color = "#55A868") +
      geom_point(data = binned_rain, aes(humidity, rain_rate), size = 3) +
      ylim(0, 1) +
      labs(x = "humidity at 9 a.m.", y = "probability of rain tomorrow") +
      theme_minimal(base_size = 13)
  })
  
  output$posteriorCode <- renderText({
    stan_call_text(input$m0, input$s0, input$m1, input$s1, prior_pd = FALSE)
  })
  
  output$diagnostics <- renderPrint({
    req(fitted_model())
    # summary() already includes Rhat and n_eff for every parameter -- this
    # avoids the rhat() name collision between rstanarm and the posterior package.
    print(summary(fitted_model())[, c("Rhat", "n_eff")])
  })
  
  ## ===== TAB 4: Prediction & classification (fast, reuses draws) =====
  
  output$predictionStatus <- renderUI({
    if (is.null(fitted_draws())) p(em("Fit a posterior model on Tab 3 first -- this tab reuses those draws."))
  })
  
  sim_prediction <- reactive({
    req(fitted_draws())
    d <- fitted_draws()
    set.seed(84735)
    log_odds <- d$`(Intercept)` + d$humidity9am * input$h_predict
    prob <- plogis(log_odds)
    tibble(log_odds = log_odds, odds = exp(log_odds), prob = prob, Y = rbinom(length(prob), 1, prob))
  })
  
  output$verdict <- renderUI({
    s <- sim_prediction()
    p_rain <- mean(s$Y == 1)
    classify_rain <- p_rain >= input$cutoff
    ci <- quantile(s$prob, c(0.1, 0.9))
    cls <- if (classify_rain) "verdict verdict-rain" else "verdict verdict-dry"
    div(class = cls,
        strong(if (classify_rain) "Carry an umbrella: classify as RAIN." else "Leave the umbrella: classify as NO RAIN."),
        br(),
        sprintf(
          paste0(
            "At %.0f%% humidity, %s of the %s posterior predictions called for rain. Since %s %s the cutoff of %s, ",
            "we classify tomorrow as %s. (The underlying chance of rain, \u03c0, has a posterior median of %s ",
            "with an 80%% credible interval of %s to %s.)"
          ),
          input$h_predict, pct(p_rain, 1), format(nrow(s), big.mark = ","), pct(p_rain, 1),
          if (classify_rain) "meets or exceeds" else "falls below", input$cutoff,
          if (classify_rain) "rain" else "no rain", pct(median(s$prob), 1), pct(ci[1], 1), pct(ci[2], 1)
        )
    )
  })
  
  output$predictionPlot <- renderPlot({
    s <- sim_prediction()
    tibble(Y = factor(s$Y, labels = c("No rain (Y = 0)", "Rain (Y = 1)"))) %>%
      count(Y) %>%
      ggplot(aes(Y, n, fill = Y)) +
      geom_col(show.legend = FALSE) +
      scale_fill_manual(values = c("#d9962b", "#2c6fbb")) +
      labs(x = NULL, y = "posterior predictions",
           title = sprintf("Posterior predictive model of tomorrow's rain at %.0f%% humidity", input$h_predict)) +
      theme_minimal(base_size = 13)
  })
  
  output$predictionCode <- renderText({
    sprintf(
      paste0(
        "posterior <- as.data.frame(rain_model)\n",
        "log_odds <- posterior$`(Intercept)` + posterior$humidity9am * %d\n",
        "prob <- exp(log_odds) / (1 + exp(log_odds))\n",
        "Y <- rbinom(length(prob), size = 1, prob = prob)\n",
        "mean(Y)   # proportion of predictions calling for rain, compare to cutoff = %s"
      ),
      input$h_predict, input$cutoff
    )
  })
  
  # Per-day posterior rain probability for all observed days (computed once per fit)
  day_probs <- reactive({
    req(fitted_draws())
    d <- fitted_draws()
    lin <- outer(d$`(Intercept)`, rep(1, n_days)) + outer(d$humidity9am, weather$humidity9am)
    colMeans(plogis(lin))
  })
  
  metrics_now <- reactive(class_metrics(day_probs(), weather$y, input$cutoff))
  
  output$metricsStatement <- renderUI({
    m <- metrics_now()
    lean <- if (input$cutoff < 0.35) "You've set a low bar for calling rain, so you catch most real rainy days but carry an umbrella a lot for nothing."
    else if (input$cutoff > 0.65) "You've set a high bar for calling rain, so you rarely carry an umbrella needlessly but you miss many real rainy days."
    else "This is a middle-of-the-road cutoff, balancing the two kinds of mistakes."
    div(class = "interp-panel",
        sprintf(
          paste0(
            "At a cutoff of %s: overall accuracy %s, sensitivity %s (of the days it actually rained, the share ",
            "we correctly called), specificity %s (of the dry days, the share we correctly called). %s Note: ",
            "accuracy alone can mislead here, since only about %s of days are rainy -- a model that ",
            "always said \"no rain\" would score %s accuracy while being useless."
          ),
          input$cutoff, pct(m$accuracy, 1), pct(m$sensitivity, 1), pct(m$specificity, 1), lean,
          pct(base_rate, 0), pct(1 - base_rate, 0)
        )
    )
  })
  
  output$confusionTable <- renderTable({
    m <- metrics_now()
    as.data.frame.matrix(m$cm) %>% rownames_to_column("Actual \\ Predicted")
  }, digits = 0)
  
  output$tradeoffPlot <- renderPlot({
    cuts <- seq(0.05, 0.95, by = 0.025)
    probs <- day_probs()
    tab <- map_dfr(cuts, function(cc) {
      m <- class_metrics(probs, weather$y, cc)
      tibble(cutoff = cc, Sensitivity = m$sensitivity, Specificity = m$specificity, `Overall accuracy` = m$accuracy)
    }) %>% pivot_longer(-cutoff, names_to = "metric", values_to = "value")
    ggplot(tab, aes(cutoff, value, color = metric)) +
      geom_line(linewidth = 1.1) +
      geom_vline(xintercept = input$cutoff, linetype = "dashed") +
      scale_color_manual(values = c(Sensitivity = "#2c6fbb", Specificity = "#d9962b", `Overall accuracy` = "grey40")) +
      labs(x = "classification cutoff", y = NULL, color = NULL,
           title = "Sensitivity and specificity trade off as the cutoff moves (dashed = your choice)") +
      theme_minimal(base_size = 13) + theme(legend.position = "top")
  })
  
  output$classificationCode <- renderText({
    sprintf("classification_summary(model = rain_model, data = weather, cutoff = %s)", input$cutoff)
  })
}

shinyApp(ui, server)

```
