---
title: "PMA Analysis"
format: html
editor: visual
---

## Libraries

```{r}
library(tidyverse)
library(bayesrules)
library(BayesFactor)
library(haven)
library(here)
library(brms)
library(janitor)
```

## Data

NHANES

```{r}
diet_behavior <- read_xpt(here("data/nhanes_data/DR1TOT_J.XPT"))
blood_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Blood.XPT"))
urine_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Urine.XPT"))
diabetes <- read_xpt(here("data/nhanes_data/2017-2018_Diabetes.XPT"))
demographics <- read_xpt(here("data/nhanes_data/2017-2018_Demographics.XPT"))
body_measures <- read_xpt(here("data/nhanes_data/BMX_J.xpt"))


```

### Subset Read-in Datasets

Subset 'diet_behavior' as 'diet'

```{r}
diet <- select(diet_behavior, SEQN, DRD360, DRD370B, DRD370BQ, DRD370Q, DRD370QQ)
```

Subset 'diabetes' as 'tiid'

```{r}
tiid <- select(diabetes, SEQN, DIQ010, DIQ170)
```

Subset 'blood_hg' as 'bhg'

```{r}
bhg <- select(blood_hg, SEQN, LBXIHG, LBDIHGSI, LBXBGE, LBXBGM)
```

Subset "urine_hg' as 'uhg'

```{r}
uhg <- select(urine_hg, SEQN, URXUHG)
```

Subset BMI

```{r}
bmi <- select(body_measures, SEQN, BMXBMI)
```

Subset Demographics

```{r}
dems <- select(demographics, SEQN, RIDAGEYR, RIAGENDR)
```

### Merge Subsets Into A Working Dataframe as 'df'

```{r}
df <- list(diet, tiid, bhg, uhg, bmi, dems)

df <- df %>% reduce(full_join, by = 'SEQN')
```

1.  Filter Dataframe df for the following:

```{r}
# Assuming your dataframe is named `nhanes_data`
df <- df %>%
  # Filter out rows where DRD370B is NA
  filter(!is.na(DRD370B)) %>%
  # Recode 1 to "Yes" and 2 to "No" for DIQ010 and DRD360
  mutate(
    DRD370B = ifelse(DRD370B == 1, "Consumes Ahi", "No Ahi")
  )
  
```

```{r}
df <- df %>% 
  filter(!is.na(LBXBGM))
```

methods

```{r}
tabyl(df$RIAGENDR)
```

```{r}
mean(df$RIDAGEYR)
sd(df$RIDAGEYR)
```

```{r}
tabyl(df$DRD370B)
```

```{r}
summary(df$LBXBGM)
```

```{r}
ggplot(data = df, aes(x = as.factor(DRD370BQ), y = LBXBGM, fill = as.factor(DRD370BQ))) + 
  labs(title = "# of times ahi waten in past 30 days with methylmercury") +
  geom_boxplot() +
  geom_jitter(size = 0.05) +
  ylim(0,20) +
  theme_minimal()
```

### Analysis

Standard boxplot

```{r}
ggplot(data = df, aes(x = DRD370B, y = LBXBGM, fill = DRD370B)) +
  geom_boxplot() +
  #geom_jitter(size = 0.001) +
  #ylim(0,5)+
  theme_minimal()

```

Log Violin plot w/ boxplots

why log?

Using **log-transformed blood mercury levels** makes the analysis more **statistically robust**, helps meet model assumptions, and allows us to interpret results in terms of **relative differences** rather than absolute changes, which is often more meaningful in environmental health studies.

**Right-Skewed Distribution**:\
Blood mercury levels are typically **highly right-skewed**, meaning most people have low levels, while a few individuals have very high concentrations. Log transformation **reduces skewness**, making the data more **approximately normal**, which is often desirable for modeling.

```{r}
ggplot(df, aes(x = DRD370B, y = log(LBXBGM), fill = DRD370B)) +
  geom_violin(alpha = 0.7, trim = FALSE) +  # Preserve full distribution
  geom_boxplot(width = 0.1, color = "black", alpha = 0.7, outlier.shape = NA) +  # Overlay boxplot
  stat_summary(fun = median, geom = "point", size = 3, color = "black") +  # Add median
  #scale_y_continuous(trans = "log10") +  # Log transformation for better scaling
  scale_fill_manual(values = c("cyan4", "darkorange")) +  # Custom colors
  labs(
    x = "Ahi Consumption",
    y = "Blood Mercury Levels (µg/L log)",
    title = "Log Distribution of Blood Mercury Levels by Ahi Consumption"
  ) +
  theme_minimal(base_size = 14) +  # Clean theme
  theme(legend.position = "top") 
  # geom_hline(yintercept = log(4.7)) +
  # annotate("text", y = log(4.71), x= 0.1, label = "HOMA-IR threshold log (4.71 μg/L)", color = "red", vjust = -4, hjust = -0.5)
```

"Consistent with the U.S. EPA reference dose, a threshold of 5.8 µg/L was used to identify participants with elevated blood mercury levels (<https://pmc.ncbi.nlm.nih.gov/articles/PMC5584810/>)."

<https://pubmed.ncbi.nlm.nih.gov/31089517/> 4.71 cut off value for insulin resistance in men

```{r}
ggplot(df, aes(x = LBXBGM, fill = DRD370B)) +
  geom_histogram(bins = 60, color = "black") +
  xlim(0,20)
```

```{r}
threshold <- 4.7 # Set the threshold for blood mercury level

df %>%
  group_by(DRD370B) %>%
  summarise(
    total = n(),
    exceeding = sum(LBXBGM > threshold, na.rm = TRUE),
    percentage_exceeding = (exceeding / total) * 100
  )
```

Perform a Bayesian t-test

```{r}
ttestBF(
  formula = LBXBGM ~ DRD370B,  # Numeric variable on left, categorical on right
  data = df
)
```

Show difference in means

```{r}
ggplot(df, aes(x = DRD370B, y = LBXBGM)) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2) +
  labs(x = "Fish Consumption (ahi)", y = "Blood Mercury Level")

```

Show difference in density

```{r}
ggplot(data = df, aes(x = LBXBGM, fill = DRD370B)) +
  geom_density(alpha = 0.7) +
  theme_bw()
```

Log Bayesian Regression Model

```{r}
log_model <- brm(LBXBGM ~ DRD370B, data = df, family = Gamma(link = "log"))
summary(log_model)
```

Plot Posterior Density of Log Difference

```{r}
log_posterior <- posterior_samples(log_model)


ggplot(log_posterior, aes(x = b_DRD370BNoAhi)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "",
    x = "Log Difference in Blood Mercury Levels (Non-Ahi vs. Ahi Consumers)",
    y = "Posterior Density of Log Difference"
  ) +
  theme_bw()

```

Post Predictive check

```{r}
brms::pp_check(log_model)
```

```{r}
pp_check(log_model, type = "dens_overlay", ndraws = 100) +
  labs(
    title = "Posterior Predictive Check",
    x = "Blood Mercury Levels (µg/L)",
    y = "Density",
    color = "Legend"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

Bayesian Regression Model

```{r}
model <- brm(LBXBGM ~ DRD370B, data = df, family = Gamma())
summary(model)
```

```{r}
ggplot(data = df, aes(x = LBXBGM, fill = DRD370B)) +
  geom_density(alpha = 0.2) +
  scale_fill_manual(values = c("cyan4", "darkorange")) +
  geom_vline(xintercept = 4.1, linetype = "dashed", color = "black", linewidth = 1) +
  annotate("text", x = 4.71, y = 0.1, label = "HOMA-IR threshold (4.71 μg/L)", color = "red", angle = 90, vjust = -4, hjust = -0.5) +
  xlim(0, 10) +
  theme_minimal()
```

```{r}
posterior <- posterior_samples(model)


ggplot(log_posterior, aes(x = b_DRD370BNoAhi)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "",
    x = "Difference in Blood Mercury Levels (Non-Ahi vs. Ahi Consumers)",
    y = "Posterior Density of Difference"
  ) +
  theme_bw()
```

```{r}
pp_check(model, type = "dens_overlay", ndraws = 100) +
  labs(
    title = "Posterior Predictive Check",
    x = "Blood Mercury Levels (µg/L)",
    y = "Density",
    color = "Legend"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

test diabetes but no luck

```{r}
# brm(DIQ010 ~ LBXBGM + DRD370B + RIDAGEYR + RIAGENDR + BMXBMI, data = df, family = bernoulli())

```
