---
title: "model choices"
format: html
editor: visual
---

```{r}
library(bayesrules)
library(here)
library(tidyverse)
```

### Alex & Fletcher

-   Modeling the average attendance at an MLB game

Prior belief

-   The average attendance is 30,000 fans, ranging between 29,500 and 30,500.
    -   sd = √a/b = 174
    -   95% of the data falls within 29,652 and 30,348

```{r}
plot_gamma(shape = 300000, rate = 1, mean = TRUE, mode = TRUE)
```

New data (Likelihood)

```{r}
y_mlb = c(31876,30500,33210,28900,35000,31050,32000,32800,29500,34000,
         36000,30000,31500,33000,31000,32500,33500,29900,34500,32200)
mlb_attendance <- data.frame(
  game = 1:length(y_mlb),
  attendance = y_mlb
)
```

```{r}
ggplot(mlb_attendance, aes(x = attendance)) +
  geom_histogram(bins = 5)
```

Posterior

```{r}
plot_gamma_poisson(shape = 30000, rate = 1, sum_y = 642936, n = 20)
```

```{r}
summarize_gamma_poisson(shape = 30000, rate = 1, sum_y = 642936, n = 20)
```

Challenge

-   Pull in some real MLB data on stadium attendance

```{r}
sea_attendance <- read_csv(here("data/sportsref_download.csv"))
```

```{r}
ggplot(sea_attendance, aes(x = Attendance)) +
  geom_density()
```

-   Try modeling average attendance with a normal-normal model

```{r}
plot_normal(mean = 30000, sd = 5000)
```

```{r}
sd(sea_attendance$Attendance)
```

```{r}
plot_normal_likelihood(y = sea_attendance$Attendance)
```

```{r}
plot_normal_normal(mean = 30000, sd = 1000, sigma = 9838,
                   y_bar = 29406, n = 162)
```

```{r}
summarize_normal_normal(mean = 30000, sd = 5000, sigma = 9838,
                   y_bar = 29406, n = 162)
```

-   Try utilizing a beta-binomial model with *stadium capacity* and *fill rate*.

```{r}
plot_beta(10, 2)
```

```{r}
plot_beta_binomial(alpha = 100, beta = 20, y = 1000, n = 2000)

```

```{r}
capacity <- 53000
sea_attendance <- sea_attendance %>%
  mutate(fill = Attendance / capacity)
```

```{r}

```

-   Which model makes the most sense and why?
