---
title: "pop_vs_soda"
format: html
editor: visual
---

```{r, message=FALSE}
library(bayesrules)
library(tidyverse)
library(janitor)
library(scales)
```

#### Import data with bayesrules package

```{r}
data("pop_vs_soda")
```

#### Bring up dataset documentation

```{r}
?pop_vs_soda
```

#### Columns of Interest

1\) region

-   region in which the state that the respondent is from falls

```{r}
unique(pop_vs_soda$region)
```

2\) pop

-   whether or not the respondent refers to fizzy cola drinks as "pop"

-   we will change the values to be "says pop" or "no pop"

```{r}
unique(pop_vs_soda$pop)
```

```{r}
pop_vs_soda <- pop_vs_soda %>%
  mutate(pop = ifelse(pop, "says pop", "no pop"))
```

```{r}
unique(pop_vs_soda$pop)
```

#### Bayes Theorem

$$
P(B \mid A) = \frac{P(B) L(B \mid A)}{P(A)} 
$$

A \<- says pop

-   Ac \<- no pop

B \<- from the south

-   Bc \<- from somewhere else

#### P(A)

These priors are given to us by the textbook from the US Census

-   Represent the united states regional distributions better than our dataset

```{r}
priors <- c(
  midwest = 0.21,
  northeast = 0.17,
  south = 0.38,
  west = 0.24
)

priors
```

```{r}
prior_probability_from_south <- 0.38
```

#### L(B\|A) = P(A\|B)

The probability that the respondent says pop given that they are from the south

```{r}
pop_vs_soda %>% 
  tabyl(pop, region) %>% 
  adorn_percentages("col")
```

```{r}
likelihoods <- pop_vs_soda %>%
  tabyl(pop, region) %>%
  adorn_percentages("col") %>%
  filter(pop == "says pop") %>%
  select(-pop) %>%
  unlist()

likelihoods
```

3\) Marginal prior probability that a person uses the word pop

```{r}
prior_probability_says_pop <- sum(priors * likelihoods)
prior_probability_says_pop
```

```{r}
pop_vs_soda %>% 
  tabyl(pop)
```

4\) Finding our posterior probability P(S\|A)

```{r}
(priors[3] * likelihoods[3]) / prior_probability_says_pop
```
