---
title: "Naive Bayes Breast Cancer"
format: html
editor: visual
---

## Dataset & Paper:

The ***Breast Cancer Wisconsin (Diagnostic)*** dataset is a renowned collection of data used extensively in machine learning and medical research. Originating from digitized images of fine needle aspirates (FNA) of breast masses, this dataset facilitates the analysis of cell nuclei characteristics to aid in the diagnosis of breast cancer. In this article, we delve into the attributes, statistics, and significance of this dataset.

[*A new classifier for breast cancer detection based on Naïve Bayesian*](https://www.sciencedirect.com/science/article/abs/pii/S0263224115002419)

### Libraries

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(e1071)
library(here)
library(ggcorrplot)
library(ggpubr)
library(janitor)
options(scipen = 999)
```

### Read in Data

```{r}
breast_cancer_data <- read_csv(here("data/wisconsin_breast_cancer_data.csv"))
```

### Correlation Matrix

Let's test the assumption that are features are independent from each other

Drop columns that we don't need in the correlation matrix

```{r}
breast_cancer_data_cor <- breast_cancer_data %>% 
  select(-id, -diagnosis, -...33)
```

Create a matrix from the dataframe

```{r}
breast_cancer_matrix <- as.matrix(breast_cancer_data_cor)
```

Calculate the correlations using `cor()`

```{r}
breast_cancer_correlations <- cor(breast_cancer_matrix)
```

Plot the correlations

```{r}
ggcorrplot(breast_cancer_correlations, type = "lower", lab = TRUE, lab_size = 1)
```

visualizing correlations between two features

```{r}
ggplot(breast_cancer_data, aes(x = radius_mean, y = perimeter_mean)) +
  geom_point() +
  geom_smooth() +
  stat_cor()
```

### Create a Naive Bayes Model

```{r}
breast_cancer_mod <- naiveBayes(diagnosis ~., data = breast_cancer_data)
```

predict diagnosis in breast cancer dataset

```{r}
breast_cancer_data <- breast_cancer_data %>% 
  mutate(predicted_diagnosis = predict(breast_cancer_mod, newdata = .))
```

create the confusion matrix

```{r}
breast_cancer_data %>% 
  tabyl(diagnosis, predicted_diagnosis) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

```{r}
breast_cancer_data <- breast_cancer_data %>% 
  mutate(predicted_diagnosis_raw = predict(breast_cancer_mod, newdata = ., type = "raw"))
```

```{r}
ggplot(breast_cancer_data, aes(x = predicted_diagnosis_raw[, "M"])) +
  geom_histogram(bins = 100)
```

```{r}
ggplot(breast_cancer_data, aes(x = predicted_diagnosis_raw[, "B"])) +
  geom_histogram(bins = 100)
```
