---
title: "Naive Bayes Assignment"
format: html
editor: visual
---

## Load Libraries

```{r, message=FALSE, warning=FALSE}
library(bayesrules)
library(tidyverse)
library(e1071)
library(titanic)
library(janitor)
```

### True vs Fake News

**Exercise 14.7 (Fake news: three predictors)** Suppose a ***new news article*** is posted online – it has a 15-word title, 6% of its words have negative associations, and its title *doesn’t* have an exclamation point. We want to know if it is fake or real

#### Read in data

```{r}
data(fake_news)
```

#### Visualization (Exploratory Data Analysis) - 2 points

Use `ggplot` to visualize the features of the data we are interested in

-   Type (fake vs real)

-   Number of words in the title (numeric value)

-   Negative associations (numeric value)

-   Exclamation point in the title (true vs false)

```{r}
ggplot(fake_news, aes(x = title_words, y = negative, color = title_has_excl)) +
  geom_point() +
  facet_wrap(~type) +
  geom_vline(xintercept = 15) +
  geom_hline(yintercept = 6)
```

#### Interpretation of Visualization - 2 points

Write a few sentences explaining whether or not this ***new news article*** is true or fake solely using your visualization above

#### Perform Naive Bayes Classification - 2 points

Based on these three features (15-word title, 6% of its words have negative associations, and its title *doesn’t* have an exclamation point), utilize naive Bayes classification to calculate the posterior probability that the article is real. Do so using `naiveBayes()` with `predict()`.

```{r}
naive_model_news <- naiveBayes(type ~ negative + title_words + title_has_excl, data = fake_news)
```

```{r}
our_article <- data.frame(negative = 6, title_words = 15, title_has_excl = "FALSE")
```

```{r}
predict(naive_model_news, newdata = our_article, type = "raw")
```

#### Break Down the Model - 6 points

Similar to the penguins example, we are going to break down the model we created above. To do this we need to find:

-   Probability(15 - word title\| article is real) using `dnorm()`

-   Probability(6% of words have negative associations \| article is real) using `dnorm()`

-   Probability(no exclamation point in title \| article is real)

    -   Probability(15 - word title\| article is fake) using `dnorm()`

    -   Probability(6% of words have negative associations \| article is fake) using `dnorm()`

    -   Probability(no exclamation point in title \| article is fake)

```{r}
naive_model_news
```

##### Real

Negative Associations

```{r}
dnorm(6, mean = 2.8, sd = 1.2)
```

Title Words

```{r}
dnorm(15, mean = 10.4, sd = 3.2)
```

```{r}
probs_true <- 0.6 * 0.009496655 * 0.04436553 * 0.97
probs_true
```

##### Fake

Negative Associations

```{r}
dnorm(6, mean = 3.6, sd = 1.5)
```

Title Words

```{r}
dnorm(15, mean = 12.3, sd = 3.7)
```

```{r}
probs_false <- 0.4 * 0.07394722 * 0.0826183 * 0.73
probs_false
```

```{r}
probs_true / (probs_false + probs_true)
```

#### Confusion Matrix

```{r}
fake_news <- fake_news %>% 
  mutate(predicted_type = predict(naive_model_news, newdata = .))
```

```{r}
fake_news %>% 
  tabyl(type, predicted_type) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

#### How can our model be improved?

```{r}
naive_model_news_all <- naiveBayes(type ~ ., data = fake_news)
```

```{r}
fake_news <- fake_news %>% 
  mutate(predicted_type_all = predict(naive_model_news_all, newdata = .))
```

```{r}
fake_news %>% 
  tabyl(type, predicted_type_all) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```
