---
title: "Homework 1"
format: html
editor: visual
---

#### Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(tidyverse)
library(haven)
library(here)
library(ggplot2)
library(janitor)
library(scales)
library(bayesrules)
library(shiny)
```

#### Read-in Datasets for this Analysis:

-   DR1TOT_J.xpt as 'diet_behavior'
-   2017-2018_Hg-Blood.XPT as 'blood_hg'
-   2017-2018_Hg-Urine.XPT as 'urine_hg'
-   2017-2018_Diabetes.XPT as 'diabetes'

```{r}
diet_behavior <- read_xpt(here("data/nhanes_data/DR1TOT_J.XPT"))
blood_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Blood.XPT"))
urine_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Urine.XPT"))
diabetes <- read_xpt(here("data/nhanes_data/2017-2018_Diabetes.XPT"))
demographics <- read_xpt(here("data/nhanes_data/2017-2018_Demographics.XPT"))
```

#### Subset Read-in Datasets

Subset 'diet_behavior' as 'diet'

```{r}
diet <- select(diet_behavior, SEQN, DRD360, DRD370B, DRD370BQ, DRD370Q, DRD370QQ)
```

Subset 'diabetes' as 'tiid'

```{r}
tiid <- select(diabetes, SEQN, DIQ010)
```

Subset 'blood_hg' as 'bhg'

```{r}
bhg <- select(blood_hg, SEQN, LBXIHG, LBDIHGSI, LBXBGE, LBXBGM)
```

Subset "urine_hg' as 'uhg'

```{r}
uhg <- select(urine_hg, SEQN, URXUHG)
```

#### Merge Subsets Into A Working Dataframe as 'df'

```{r}
df <- list(diet, tiid, bhg, uhg)

df <- df %>% reduce(full_join, by = 'SEQN')
```

Filter Dataframe

```{r}
# Assuming your dataframe is named `nhanes_data`
df <- df %>%
  # Filter out rows where DIQ010 or DRD360 are NA
  filter(!is.na(DIQ010), !is.na(DRD370B)) %>%
  # Keep only rows where DIQ010 and DRD360 are 1 or 2
  filter(DIQ010 %in% c(1, 2), DRD370B %in% c(1, 2)) %>%
  # Recode 1 to "Yes" and 2 to "No" for DIQ010 and DRD360
  mutate(
    DIQ010 = ifelse(DIQ010 == 1, "Has Diabetes", "No Diabetes"),
    DRD370B = ifelse(DRD370B == 1, "Consumes Ahi", "No Ahi")
  )
  
```

#### Bayes Theorem

$$
P(B \mid A) = \frac{P(B) L(B \mid A)}{P(A)} 
$$

A \<- consumes ahi

-   Ac \<- no ahi

B \<- has diabetes

-   Bc \<- no diabetes

#### P(B)

Prior Probability that a person has diabetes

-   [check national average outside dataset](https://www.cdc.gov/diabetes/php/data-research/index.html#:~:text=Prevalence%20of%20both%20diagnosed%20and%20undiagnosed%20diabetes&text=Among%20the%20U.S.%20population%20overall,or%20older%20(Table%201a).)

```{r}
df %>% 
  tabyl(DIQ010) 
```

```{r}
prior_probability_has_diabetes = 0.13
```

#### L(B\|A) = P(A\|B)

The probability of a person consuming ahi given that they have diabetes

```{r}
df %>% 
  tabyl(DIQ010, DRD370B) %>% 
  adorn_percentages("row")
```

```{r}
nhanes_percent <- df %>%
  tabyl(DIQ010, DRD370B) %>%
  adorn_percentages("row")

# Step 2: Convert the wide format into a long format
nhanes_long <- nhanes_percent %>%
  pivot_longer(cols = -DIQ010, names_to = "DRD370B", values_to = "percentage")


# Step 3: Create the bar plot with percentages
ggplot(nhanes_long, aes(x = DIQ010, y = percentage, fill = DRD370B)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_text(aes(label = percent(percentage, accuracy = 0.1)), 
            position = position_fill(vjust = 0.5)) +
  scale_y_continuous(labels = percent_format()) +
  labs(y = "Percentage", x = "Diabetes", fill = "Ahi Consumption") +
  theme_minimal()
```

```{r}
likelihood_diabetes_ahi = 0.36
 

```

#### P(A)

The probability that a person consumes ahi

```{r}
df %>% 
  tabyl(DRD370B)
```

```{r}
probability_consumes_ahi <- 0.36
```

#### P(B\|A)

The posterior probability that a person has diabetes given that they eat ahi

```{r}
(prior_probability_has_diabetes * likelihood_diabetes_ahi) / probability_consumes_ahi
```

#### Other nhanes variables

```{r}
demographics <- demographics %>% 
  select(SEQN, RIDRETH1, RIAGENDR, RIDAGEYR)
```

```{r}
df2 <- left_join(df, demographics, by = "SEQN")
```

```{r}
ggplot(df2, aes(x = RIDAGEYR, fill = DIQ010)) +
  geom_density(alpha = 0.5)
```

```{r}
df2 <- df2 %>% 
  mutate(age_bracket = ifelse(RIDAGEYR > 40, "old", "young"))
```

```{r}
table(df2$age_bracket)
```

```{r}
nhanes_age_percent <- df2 %>%
  tabyl(DIQ010, age_bracket) %>%
  adorn_percentages("row")

# Step 2: Convert the wide format into a long format
nhanes_age_long <- nhanes_age_percent %>%
  pivot_longer(cols = -DIQ010, names_to = "age_bracket", values_to = "percentage")


# Step 3: Create the bar plot with percentages
ggplot(nhanes_age_long, aes(x = DIQ010, y = percentage, fill = age_bracket)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_text(aes(label = percent(percentage, accuracy = 0.1)), 
            position = position_fill(vjust = 0.5)) +
  scale_y_continuous(labels = percent_format()) +
  labs(y = "Percentage", x = "Diabetes", fill = "Age Bracket") +
  theme_minimal()
```

Bayes Theorem

P(B)

```{r}
prior_probability_has_diabetes <- 0.13
```

L(B\|A) = P(A\|B)

```{r}
df2 %>% 
  tabyl(DIQ010, age_bracket) %>% 
  adorn_percentages("row")
```

```{r}
likelihood_old_given_diabetes <- 0.96
```

P(A)

```{r}
df2 %>% 
  tabyl(age_bracket) 
```

```{r}
probability_old <- 0.53
```

P(B\|A)

```{r}
(prior_probability_has_diabetes * likelihood_old_given_diabetes) / probability_old
```

### TESTS

```{r}
df %>% 
  tabyl('DIQ010')

prior_probabilioty_has_diabetes <- 0.13
```

P(A)

```{r}
probability_over_40 <- mean(demographics$RIDAGEYR > 40, na.rm = TRUE)
probability_over_40
```

```{r}
demographics %>% 
  mutate(age_bracket = ifelse(RIDAGEYR > 40, "old", "young")) %>% 
  tabyl(age_bracket)
```

P(A\|B)

```{r}
joined <- inner_join(demographics, diabetes, by = "SEQN")

probabability_over_40_and_diabetes <- mean(joined$RIDAGEYR[joined$DIQ010 == 1] > 40, na.rm = TRUE)
probabability_over_40_and_diabetes

```

Probability Via Bayes

```{r}
Posterior_prob <- (probabability_over_40_and_diabetes * prior_probabilioty_has_diabetes) / probability_over_40
Posterior_prob
```
